### 2016-12-30 paper/paper.tex
Major work done. Fixed bug in multi-tag scenario where the error only used a single tag. Committing some previous work on the paper and some miscellaneous code maintanance.

\documentclass[11pt]{article}
%Gummi|065|=)
\title{\textbf{AprilTrack: A Particle-Filter based AprilTag Tracker}}
\author{Daniel Pfrommer}
\date{}

\begin{document}

\maketitle

\begin{abstract}

	The AprilTrack algorithm proposed in this paper seeks to bridge two existing fields of research in computer vision literature: fiducial detection and monocular object tracking. With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth position data for mobile robots, there has been a considerable interest in designing both robust and blur-resistant fiducials. This paper attempts to improve on existing fiducials such as the AprilTag marker by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom. This novel particle-filter based algorithm is evaluated on several datasets involving both ego-motion and tag-based-motion and is compared to standalone tag detection algorithms.

\end{abstract}

\section{Introduction}

	The majority of previous literature in fiducial detection and monocular object tracking has been focused single-frame fiducial detection 3

\section{Feedback}

\end{document}

### 2016-12-30 AprilTrack.m

@@ -6,35 +6,35 @@ classdef AprilTrack < TagSource
         % A cell array of motion models
         motionModels 
         
-        params
+        tagParams
     end
     
     methods
-        function obj = AprilTrack(params)
+        function obj = AprilTrack(tagParams)
             obj.motionModels = {};
             
-            obj.params = params;
+            obj.tagParams = tagParams;
             
             % Pre-calculate the projection coordinates                        
-            coordinates = zeros([params.patchSize 2]);
+            coordinates = zeros([tagParams.patchSize 2]);
             
-            for i=1:params.patchSize(2)
-                for j=1:params.patchSize(1)
-                    idx = [i / params.patchSize(2) - 0.5; ...
-                           j / params.patchSize(1) - 0.5];
+            for i=1:tagParams.patchSize(2)
+                for j=1:tagParams.patchSize(1)
+                    idx = [i / tagParams.patchSize(2) - 0.5; ...
+                           j / tagParams.patchSize(1) - 0.5];
 
                     coordinates(j, i, :) = idx;
                 end
             end
             
-            obj.params.coords = reshape(permute(coordinates, [3, 2, 1]), ...
-                                        2, params.patchSize(2) * params.patchSize(1));
-            obj.params.coords = [obj.params.coords; ones([1 size(obj.params.coords, 2)])];
+            obj.tagParams.coords = reshape(permute(coordinates, [3, 2, 1]), ...
+                                        2, tagParams.patchSize(2) * tagParams.patchSize(1));
+            obj.tagParams.coords = [obj.tagParams.coords; ones([1 size(obj.tagParams.coords, 2)])];
         end
         
         function addMotionModel(this, mm)
-            mm.setTagParams(this.params);
             this.motionModels{length(this.motionModels) + 1} = mm;
+            mm.setTagParams(this.tagParams);
         end
         
         function tags = process(this, img, detector_tags)
@@ -53,9 +53,9 @@ classdef AprilTrack < TagSource
                 tags = [tags mm.process(img, tagMap)];
             end
         end
-        function debug(this, fig1, fig2, fig3)
+        function debug(this, fig1, fig2, fig3, fig4)
             if length(this.motionModels) > 0
-                this.motionModels{1}.debug(fig1, fig2, fig3);
+                this.motionModels{1}.debug(fig1, fig2, fig3, fig4);
             end
         end
     end

### 2016-12-30 ImagesSource.m

@@ -3,26 +3,31 @@ classdef ImagesSource < FrameSource
     %   Detailed explanation goes here
     
     properties
-        dir
         files
+        index
+        
+        skip
         
         useUndistort
         camParams
     end
     
     methods
-        function obj = ImagesSource(directory, camParams)
+        function obj = ImagesSource(directory, skip, camParams)
             files = dir(directory);
             
             obj.files = {};
-            obj.dir = directory;
+            obj.index = 1;
+            obj.skip = skip;
+            
+            
             i = 1;
             c = 1;
             while i < length(files);
                 file = files(i).name;
                 if file(1) == '.'
                 else
-                    obj.files{c} = file;
+                    obj.files{c} = fullfile(directory, file);
                     c = c + 1;
                 end
                 i = i + 1;
@@ -39,9 +44,9 @@ classdef ImagesSource < FrameSource
         end
         
         function img = readImage(this)
-            f = fullfile(this.dir, this.files{1});
-            this.files(:, 1) = [];
-
+            f = this.files{this.index};
+            this.index = this.index + this.skip;
+            
             img = imread(f);
             if ndims(img) > 2
                 img = rgb2gray(img);

### 2016-12-30 MotionModel.m

@@ -38,6 +38,8 @@ classdef MotionModel < handle
             obj.visibleTags = [];
         end
 
+        % Will be called by the motion model to pass
+        % along the coordinates
         function setTagParams(this, tagParams)
             this.tagParams = tagParams;
         end
@@ -81,14 +83,14 @@ classdef MotionModel < handle
         end
         
         function loadTags(this, file)
-            [ids, sizes, posX, posY, posZ, rotX, rotY, rotZ, rotW] = ...
-                textread(file, '%f %f %f %f %f %f %f %f %f');
-
+            [ids, sizes, border, posX, posY, posZ, rotX, rotY, rotZ, rotW] = ...
+                textread(file, '%f %f %f %f %f %f %f %f %f %f', 'commentstyle', 'shell');
+            
             for i=1:length(ids)
                 tag(1).id = ids(i);
                 tag(1).color = 'r';
                 tag(1).size = 0.001 * [sizes(i) sizes(i)];
-                tag(1).border = [0.03 0.03];
+                tag(1).border = 0.001 * [border(i) border(i)];
 
                 pos = (0.001 .* [posX(i) posY(i) posZ(i)])';
                 rot = vrrotvec_to_quat([rotX(i) rotY(i) rotZ(i) rotW(i)])';
@@ -131,7 +133,8 @@ classdef MotionModel < handle
                     this.modelledTags{i}.refPatch = ... 
                         extract_patch(this.tagParams.patchSize, ...
                                       this.tagParams.coords, ...
-                                      img, tag);                    
+                                      img, tag);
+                    % make sure we note that the tag is visible
                     this.visibleTags(i) = 1;
                 end
             end
@@ -160,8 +163,7 @@ classdef MotionModel < handle
             tic();
 
             this.particles = q_smpl(this.particles, this.weights, 1, ...
-                                    this.params.process_noise, ...
-                                    this.params.k, this.params.alpha);
+                                    this.params.process_noise);
 
             fprintf('; Took: %f\n', toc());
 
@@ -189,25 +191,18 @@ classdef MotionModel < handle
             
             
             % Convert the measurements to weights
+            this.measurements = Z;
             W = this.transformMeasurements(Z);
             % this.measurements is really just the transformed weights
-            this.measurements = W;
             
             % Update the old weights
-            % using the formula w' = w * p(x'|x)/q(x'|x,y)
-            % where q(x') = p(x') * 1/(k + alpha*w)
[…]

### 2016-12-30 TagDetector.m

@@ -35,12 +35,18 @@ classdef TagDetector < TagSource
 
                 H = homography_solve(pin, corners);
                 [R, T] = homography_extract_pose(this.K, H);
+                r.id
                 
-                rot = rotm_to_quat(R')';
-                if r.id == 0
-                    -T
-                    rot
-                end
+                H_0 =  [0.5417    0.0623    0.6447;
+                        0.0138    0.5204    0.1269;
+                        0.0001    0.0001    0.0009];
+
+                pix_center = H(:,3) ./ H(3,3);
+                
+                T_rel = inv(H_0) * pix_center;
+                T_rel = T_rel ./ T_rel(3);
+                T_rel * 1000
+                %rot = rotm_to_quat(R')';
                 
                 %}
                 

### 2016-12-30 algorithm.m

@@ -7,12 +7,21 @@ function [ M ] = algorithm(tracker, detector, images, record)
     fig2 = sfigure(2);
     fig3 = sfigure(3);
     fig4 = sfigure(4);
+    fig5 = sfigure(5);
     
     figure(1);
     
     disp('Entering main loop');
     
+    counter = 1;
+    
     while images.hasImage()
+        
+        counter = counter + 1;
+        if mod(counter, 10) == 0
+            clc;
+        end
+        
         fprintf('------------------------------\n');      
         fprintf(':: Reading image\n'); 
         tic();
@@ -30,6 +39,7 @@ function [ M ] = algorithm(tracker, detector, images, record)
         clf(fig2);
         clf(fig3);
         clf(fig4);
+        clf(fig5);
         
         fprintf('// Took %f\n', toc());
         fprintf(':: Displaying result\n');
@@ -42,7 +52,7 @@ function [ M ] = algorithm(tracker, detector, images, record)
         
         % project the tags, will be stored in the
         % tags array
-        tags = project_tags(tracker.params.K, tags);
+        tags = project_tags(tracker.tagParams.K, tags);
         
         %drawTags(detector_tags, 'symbol', 'x');
         %drawTags(reproj_tags, 'symbol', 'o');
@@ -54,7 +64,7 @@ function [ M ] = algorithm(tracker, detector, images, record)
         fprintf(':: Displaying debug stuff\n');
         tic();
 
-        tracker.debug(fig2, fig3, fig4);
+        tracker.debug(fig2, fig3, fig4, fig5);
         sfigure(1);
 
         fprintf('// Took %f\n', toc());
@@ -72,7 +82,6 @@ function [ M ] = algorithm(tracker, detector, images, record)
 end
 
 function drawTags(tags, varargin)
-    disp('foo');
     for i=1:length(tags)
         drawTag(tags{i}, varargin);
     end
@@ -80,7 +89,7 @@ end
 
 % Draws a (projected!) tag
 function drawTag(tag, vars)
-    symbol = 'x';
+    symbol = '.-';
     for i=1:length(vars)
         if strcmp(vars{i},'symbol')
             symbol = vars{i + 1};

### 2016-12-30 config/hard_seq_config.m

@@ -1,0 +1,40 @@
+function [tracker, detector, frames] = setup_hard_seq(skip_rate)
+    frames = ImagesSource('../data/hard_seq/images', skip_rate);
+
+    % Image parameters
+    K = [568.8885   0           959.1503;
+         0          784.9941    539.7587;
+         0          0           1.0000];
+
+    % Setup some detector stuff
+    detector = TagDetector(K);
+    trackParams(1).K = K;
+    
+    % Patch size
+    trackParams(1).patchSize = [32 32];
+    
+    % Create the tracker
+    tracker = AprilTrack(trackParams);
+
+    % Add a motion model
+    mmParams(1).err_discard_threshold = 1;
+    mmParams(1).num_particles = 5000;
+    mmParams(1).process_noise = [0.04 0.04 0.04 ...
+                                 0.04 0.04 0.04 ...
+                                 0 0 0 ...
+                                 0 0 0];
+
+    % For measurement error --> weight conversion
+    % where weight = e^(-lambda * measurement)
+    mmParams(1).lambda = 10;
+
+    model = MotionModel(mmParams, @transform_ego);
+
+    model.loadTags('../data/hard_seq/tags.txt');
+
+    initial = [1.1029; 0.2798; -0.3398;  0.499; 0.499; -0.501; 0.501; 0; 0; 0; 0; 0; 0];
+    
+    model.initializeParticlesTo(initial);
+    
+    tracker.addMotionModel(model);
+end

### 2016-12-30 config/movcam1_config.m

@@ -1,0 +1,43 @@
+function [tracker, detector, frames] = movcam1_config(skip_rate)
+    frames = ImagesSource('../data/movcam1/images', skip_rate);
+
+    % Image parameters
+    K = [1885         0         946.8;
+         0            1881.3    540.8;
+         0            0         1.0];
+
+
+    % Setup some detector stuff
+    detector = TagDetector(K);
+    trackParams(1).K = K;
+    
+    % Patch size
+    trackParams(1).patchSize = [32 32];
+    
+    % Create the tracker
+    tracker = AprilTrack(trackParams);
+
+    % Add a motion model
+    mmParams(1).err_discard_threshold = 1;
+    mmParams(1).num_particles = 3000;
+    mmParams(1).process_noise = [0.04 0.04 0.04 ...
+                                 0.04 0.04 0.04 ...
+                                 0 0 0 ...
+                                 0 0 0];
+
+    % For measurement error --> weight conversion
+    % where weight = e^(-lambda * measurement)
+    mmParams(1).lambda = 10;
+
+    model = MotionModel(mmParams, @transform_ego);
+
+    model.loadTags('../data/movcam_tags.txt');
+    
+    initial = [-0.424; -0.7353; 3.4589; ...
+               0.9791; 0.1296; -0.1559; -0.0135; ...
+               0; 0; 0;  0; 0; 0];
+    
+    model.initializeParticlesTo(initial);
+    
+    tracker.addMotionModel(model);
+end

### 2016-12-30 config/test_video_config.m

@@ -1,0 +1,50 @@
+function [tracker, detector, frames] = test_video_config()
+    frames = ImagesSource('../data/test/images');
+
+    % Image parameters
+    K = [1002.6   0           540.9;
+         0          1007.0    355.4;
+         0          0         1.0];
+
+    % Setup some detector stuff
+    detector = TagDetector(K);
+    trackParams(1).K = K;
+    
+    % Patch size
+    trackParams(1).patchSize = [64 64];
+    
+    % Create the tracker
+    tracker = AprilTrack(trackParams);
+
+    % Add a motion model
+    mmParams(1).err_discard_threshold = 0.9;
+    mmParams(1).num_particles = 3000;
+    mmParams(1).process_noise = [0.01 0.01 0.01 ...
+                                 0.04 0.04 0.04 ...
+                                 0.02 0.02 0.02 ...
+                                 0 0 0];
+
+    % noise = 1 / (k + alpha * w) * process_noise * randn
+    % Allows for particles of higher/lower weight to have
+    % more noise in the propagation step
+
+    % ATM, we resample after every pass, so don't bother...
+    mmParams(1).k = 1;
+    mmParams(1).alpha = 0;
+
+    % For measurement error --> weight conversion
+    % where weight = e^(-lambda * measurement)
+    mmParams(1).lambda = 9;
+
+    model = MotionModel(mmParams, @tag_transform);
+
+    model.loadTags('../data/test/tags.txt');
+    
+    initial = [0.1862224; -0.0067017; 0.696942; ...
+               0.991367; -0.124779; -0.014736; 0.037472;...
+               0; 0; 0;  0; 0; 0];
+    
+    model.initializeParticlesTo(initial);
+    
+    tracker.addMotionModel(model);
+end

### 2016-12-30 run.m

@@ -1,16 +1,22 @@
-function [  ] = run(config, shouldRecord)
+function [ M ] = run(config, shouldRecord)
     tracker = [];
     detector = [];
     frames = [];
     M = [];
     
+    
+    % The parameter argument represents the number of frames to skip
+    % so zero means always stay on the first frame
+    
     switch config
         case 'hard_seq'
-            [tracker, detector, frames] = setup_hard_seq();
+            [tracker, detector, frames] = hard_seq_config(1);
         case 'test_video'
-            [tracker, detector, frames] = setup_test_video();
+            [tracker, detector, frames] = test_video_config();
         case 'movcam1'
-            [tracker, detector, frames] = setup_movcam1();
+            [tracker, detector, frames] = movcam1_config(1);
+        case 'movcam1-still'
+            [tracker, detector, frames] = movcam1_config(0);
         otherwise
             return;
     end
@@ -20,180 +26,4 @@ function [  ] = run(config, shouldRecord)
     else
         algorithm(tracker, detector, frames, false);
     end
-end
-
-function [tracker, detector, frames] = setup_hard_seq()
-    frames = ImagesSource('../data/hard_seq/images');
-
-    % Image parameters
-    K = [568.8885   0           959.1503;
-         0          784.9941    539.7587;
-         0          0           1.0000];
-
-    % Setup some detector stuff
-    detector = TagDetector(K);
-    trackParams(1).K = K;
-    
-    % Patch size
-    trackParams(1).patchSize = [64 64];
-    
-    % Create the tracker
-    tracker = AprilTrack(trackParams);
-
-    % Add a motion model
-    mmParams(1).err_discard_threshold = 0.9;
-    mmParams(1).num_particles = 1000;
-    mmParams(1).process_noise = [0.03 0.03 0.03 ...
-                                 0.01 0.01 0.01 ...
-                                 0 0 0 ...
-                                 0 0 0];
-
-    % noise = 1 / (k + alpha * w)
-    % Allows for particles of higher/lower weight to have
-    % more noise
-    mmParams(1).k = 1;
-    mmParams(1).alpha = 0;
-
-    % For measurement error --> weight conversion
-    % where weight = e^(-lambda * measurement)
-    mmParams(1).lambda = 10;
[…]

### 2016-12-30 save_images.m

@@ -1,4 +1,12 @@
 function save_images(input, output_dir)
+    for c=1:length(input)
+        img = input(c).cdata;
+        file = sprintf('%s/%04d.png', output_dir, c);
+        imwrite(img, file);
+    end
+end
+%{
+function save_images(input, output_dir)
     c = 1;
     while input.hasImage()
         img = input.readImage();
@@ -7,3 +15,4 @@ function save_images(input, output_dir)
         c = c + 1;
     end
 end
+%}

### 2016-12-30 setup.m

@@ -1,4 +1,5 @@
 addpath 'apriltag'
 addpath 'utils'
+addpath 'config'
 
 warning('off', 'Images:initSize:adjustingMag');

### 2016-12-30 utils/homography_from_state.m

@@ -1,0 +1,15 @@
+function [ H ] = homography_from_state(K, tag)
+
+    R = quat_to_rotm(tag.state(4:7));
+    T = tag.state(1:3);
+    H = K * [R(:, 1:2) T];
+    % Create the homography
+    
+    % Project the coords
+    s = tag.size + tag.border;
+    S = [s(1) 0 0; ...
+         0 s(2) 0; ...
+         0 0    1];
+    % Modify H to scale by size first
+    H = H * S;
+end

### 2016-12-30 utils/measure_patch_error.m

@@ -1,36 +1,36 @@
-%{ 
 % Use squared error
+%{
 function [ err ] = measure_patch_error(patchA, patchB)
-    if (size(patchA) ~= size(patchB) || ...
-        (size(patchA,1) == 0 || size(patchA,2) == 0) ||
-        (size(patchA,1) == 0 || size(patchA,2) == 0) )
+    if ((size(patchA, 1) ~= size(patchB,1)) || ...
+        (size(patchA, 2) ~= size(patchB,2)) || ...            
+        (size(patchA,1) == 0 || size(patchA,2) == 0))
         err = 1;
         return;
     else
-        a = patchA(:);
-        b = patchB(:);
-        
-        diff = double(a - b) / 255;
-        
-        err = sum(diff.^2) / length(diff);
+        [M, N] = size(patchA);
+        diff = double(patchA) - double(patchB);
+        %err = sum(sum(diff) .* diff)) / (M * N);        
+        % Divide by 255^2 to get the error from 0-1
+        err = sum(sum(diff .* diff)) / (M * N * 255 * 255);
+        return;
     end
 end
 %}
 
-
 % Use correlation
+%%{
 function [ err ] = measure_patch_error(patchA, patchB)
     if (size(patchA) ~= size(patchB))
         err = 1;
         return;
     else
-        a = int16(patchA);
-        b = int16(patchB);
+        a = double(patchA);
+        b = double(patchB);
 
-        err = abs(corr2(a, b));
-        err = 1 - err;
-        if err < 0 || isnan(err) % Some crazy value, like -Inf
-            err = 0;
+        err = 1 - max(corr2(a, b),0);
+        if err > 1 || err < 0 || isnan(err) % Some crazy value, like -Inf, Inf, NaN
+            err = 1;
         end
 
 end
+%}

### 2016-12-30 utils/q_smpl.m

@@ -9,7 +9,7 @@ function y = q_smpl(x, w, deltaT, r, k, alpha)
     for i=1:size(x, 2)
         xp = x(:, i);
         % Add some noise        
-        noise = 1 / (k + alpha * w(i)) * (r .* randn(1, 12))';
+        noise = (r .* randn(1, 12))';
         xp(:) = xp(:) + [noise(1:3); 1; 0; 0; 0; noise(7:12)];
         % Special case for the quaternion
         xp(4:7) = qmult(x(4:7, i)', rotvec_to_quat(noise(4:6)'))';

### 2016-12-30 utils/transform_ego.m

@@ -1,0 +1,10 @@
+function newPose = transform_ego(tagPose, x)
+    % Local -> World transform
+    WL = [[quat_to_rotm(tagPose(4:7)') [tagPose(1); tagPose(2); tagPose(3)]]; [0 0 0 1]];
+    % World -> Camera transform
+    CW = [[quat_to_rotm(x(4:7)') [x(1); x(2); x(3)]]; [0 0 0 1]];
+    T = CW * WL;
+    newPose = tagPose;
+    newPose(1:3) = [T(1, 4) T(2, 4) T(3, 4)];
+    newPose(4:7) = rotm_to_quat(T(1:3, 1:3));
+end

### 2016-12-30 utils/transform_tag.m

@@ -1,0 +1,5 @@
+function transState = tag_transform(state, x)
+    transState = state;
+    transState(1:3) =  x(1:3) + state(1:3);
+    transState(4:7) = qmult(x(4:7)', state(4:7)')';
+end

### 2017-01-02 FusedTagSource.m
Midway through paper

@@ -1,56 +1,0 @@
-classdef FusedTagSource < TagSource
-    %FUSEDTAGSOURCE Combines both tracking and detecting
-   
-    properties
-        detector
-        tracker
-    end
-    
-    methods
-        function obj = FusedTagSource(tagSize, K, np)
-            obj.detector = TagDetector(tagSize, K, np);
-            obj.tracker = TagTracker(np);
-        end
-        
-        function tags = process(this, img)
-            %%{
-            tracker_tags = this.tracker.process(img);
-            if size(tracker_tags, 2) > 0
-                tags = tracker_tags;
-            end
-            %}
-            detector_tags = this.detector.process(img);
-            if size(detector_tags, 2) > 0
-                tags = detector_tags;
-                this.tracker.set(tags);
-            end
-            
-            tags = merge(detector_tags, tracker_tags);
-        end
-
-    end
-    
-end
-
-
-% A function which will merge two sets
-% of tags, preferring tags in tagA over tagB
-function tags = merge(tagsA, tagsB)
-    tags = tagsA;
-
-    for i=1:length(tagsB)
-        t = tagsB{i};
-        
-        found = false;
-        for j=1:length(tagsA)
-            if tagsA{j}.id == tagsB{i}.id
-                found = true;
-                break;
-            end
-        end
-        
-        if ~found
-            tags{length(tags) + 1} = t;
-        end
-    end
-end

### 2017-01-02 ImagesSource.m

@@ -23,7 +23,7 @@ classdef ImagesSource < FrameSource
             
             i = 1;
             c = 1;
-            while i < length(files);
+            while i <= length(files);
                 file = files(i).name;
                 if file(1) == '.'
                 else
@@ -40,11 +40,13 @@ classdef ImagesSource < FrameSource
         end
         
         function hasImage = hasImage(this)
-            hasImage = (length(this.files) > 0);
+            hasImage = (length(this.files) >= this.index);
         end
         
         function img = readImage(this)
             f = this.files{this.index};
+            disp('Reading from:');
+            disp(f);
             this.index = this.index + this.skip;
             
             img = imread(f);

### 2017-01-02 MotionModel.m

@@ -195,8 +195,9 @@ classdef MotionModel < handle
             W = this.transformMeasurements(Z);
             % this.measurements is really just the transformed weights
             
-            % Update the old weights
-            this.weights = this.weights .* W;
+            % Update the old weights. We don't need to multiply as we just
+            % resampled
+            this.weights = W;
             
             % Finally, normalize the weights again
             weightSum = sum(this.weights);
@@ -268,6 +269,7 @@ classdef MotionModel < handle
         % Converts measurements to weights
         function W = transformMeasurements(this, Z) 
             W = exp(-this.params.lambda * Z);
+            %W = this.params.lambda * Z;
         end
 
         function tags = transformTags(this, x, detectedTagMap)

### 2017-01-02 algorithm.m

@@ -1,5 +1,8 @@
-function [ M ] = algorithm(tracker, detector, images, record)
-    M = [];
+function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, record)
+    imgs = [];
+    corners = [];
+    pos = [];
+    rot = [];
     
     disp('Initializing figures');
     
@@ -13,7 +16,7 @@ function [ M ] = algorithm(tracker, detector, images, record)
     
     disp('Entering main loop');
     
-    counter = 1;
+    counter = 0;
     
     while images.hasImage()
         
@@ -54,6 +57,9 @@ function [ M ] = algorithm(tracker, detector, images, record)
         % tags array
         tags = project_tags(tracker.tagParams.K, tags);
         
+        pos = [pos; tags{1}.state'];
+        corners = [corners; reshape(tags{1}.corners', [1, 8])];
+        
         %drawTags(detector_tags, 'symbol', 'x');
         %drawTags(reproj_tags, 'symbol', 'o');
         drawTags(tags);
@@ -74,7 +80,7 @@ function [ M ] = algorithm(tracker, detector, images, record)
             tic();
             fprintf('Getting frame');
 
-            M = [ M getframe(fig1) ];
+            imgs = [ imgs getframe(fig1) ];
 
             disp('-- '), disp(toc());
         end

### 2017-01-02 config/hard_seq_config.m

@@ -18,9 +18,9 @@ function [tracker, detector, frames] = setup_hard_seq(skip_rate)
 
     % Add a motion model
     mmParams(1).err_discard_threshold = 1;
-    mmParams(1).num_particles = 5000;
-    mmParams(1).process_noise = [0.04 0.04 0.04 ...
-                                 0.04 0.04 0.04 ...
+    mmParams(1).num_particles = 6000;
+    mmParams(1).process_noise = [0.05 0.05 0.05 ...
+                                 0.05 0.05 0.05 ...
                                  0 0 0 ...
                                  0 0 0];
 

### 2017-01-02 config/movcam1_config.m

@@ -20,10 +20,10 @@ function [tracker, detector, frames] = movcam1_config(skip_rate)
     % Add a motion model
     mmParams(1).err_discard_threshold = 1;
     mmParams(1).num_particles = 3000;
-    mmParams(1).process_noise = [0.04 0.04 0.04 ...
-                                 0.04 0.04 0.04 ...
-                                 0 0 0 ...
-                                 0 0 0];
+    mmParams(1).process_noise = [0.01 0.01 0.01 ...
+                                 0.01 0.01 0.01 ...
+                                 0.01 0.01 0.01 ...
+                                 0.01 0.01 0.01];
 
     % For measurement error --> weight conversion
     % where weight = e^(-lambda * measurement)

### 2017-01-02 config/test_video_config.m

@@ -1,5 +1,5 @@
 function [tracker, detector, frames] = test_video_config()
-    frames = ImagesSource('../data/test/images');
+    frames = ImagesSource('../data/test/images', 1);
 
     % Image parameters
     K = [1002.6   0           540.9;
@@ -18,11 +18,11 @@ function [tracker, detector, frames] = test_video_config()
 
     % Add a motion model
     mmParams(1).err_discard_threshold = 0.9;
-    mmParams(1).num_particles = 3000;
+    mmParams(1).num_particles = 6000;
     mmParams(1).process_noise = [0.01 0.01 0.01 ...
-                                 0.04 0.04 0.04 ...
+                                 0.05 0.05 0.05 ...
                                  0.02 0.02 0.02 ...
-                                 0 0 0];
+                                 0.00 0.00 0.00];
 
     % noise = 1 / (k + alpha * w) * process_noise * randn
     % Allows for particles of higher/lower weight to have
@@ -36,7 +36,7 @@ function [tracker, detector, frames] = test_video_config()
     % where weight = e^(-lambda * measurement)
     mmParams(1).lambda = 9;
 
-    model = MotionModel(mmParams, @tag_transform);
+    model = MotionModel(mmParams, @transform_tag);
 
     model.loadTags('../data/test/tags.txt');
     

### 2017-01-02 run.m

@@ -1,9 +1,11 @@
-function [ M ] = run(config, shouldRecord)
+function [ imgs, corners, pos, rot ] = run(config, gt, shouldRecord)
     tracker = [];
     detector = [];
     frames = [];
-    M = [];
-    
+    imgs = [];
+    corners = [];
+    pos = [];
+    rot = [];
     
     % The parameter argument represents the number of frames to skip
     % so zero means always stay on the first frame
@@ -22,7 +24,7 @@ function [ M ] = run(config, shouldRecord)
     end
     
     if (exist('shouldRecord', 'var') && shouldRecord)
-        M = algorithm(tracker, detector, frames, true);
+        [imgs, corners, pos, rot] = algorithm(tracker, detector, frames, gt, true);
     else
         algorithm(tracker, detector, frames, false);
     end

### 2017-01-02 paper/paper.tex
Paper almost done

@@
-\documentclass[11pt]{article}
-%Gummi|065|=)
+\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
+
+\usepackage{graphicx}
+\usepackage{amsmath,amsfonts,amssymb}
+
+\graphicspath{ {figures/} }
+
+\renewcommand{\vec}[1]{\boldsymbol{#1}}
+
 \title{\textbf{AprilTrack: A Particle-Filter based AprilTag Tracker}}
@@
 	The AprilTrack algorithm proposed in this paper seeks to bridge two existing fields of research in computer vision literature: fiducial detection and monocular object tracking.
-With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth position data for mobile robots, there has been a considerable interest in designing both robust and blur-resistant fiducials.
-This paper attempts to improve on existing fiducials such as the AprilTag marker by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom.
-This novel particle-filter based algorithm is evaluated on several datasets involving both ego-motion and tag-based-motion and is compared to standalone tag detection algorithms.
+With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth position data for mobile robots, there has been a considerable interest in designing robust and blur-resistant fiducials.
+Rather than devising a new fiducial, This paper attempts to improve on existing fiducials such as the popular AprilTag marker by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom, allowing for tag-based localization in situations where still frame based methods would be unable to do so.
+This novel particle-filter based algorithm is evaluated on several datasets involving both ego-motion and tag-based-motion and is compared to standalone tag detection algorithms and ground-truth tag position data.
 
@@
 
-	The majority of previous literature in fiducial detection and monocular object tracking has been focused single-frame fiducial detection 3
+Visual tracking of fiducial markers (or fiducials) plays an important role in many robotics and computer vision related fields, including augmented and virtual reality, visual odometry, and object tracking.
+Due to these scenarios taking place in tructured indoors environments where illumination is often limited, long exposure times to compensate for the reduced lighting produce blurred images, especially when taken from a mobile platform such as quadrotor.
 
-\section{Feedback}
+ The majority of previous literature in fiducial detection and monocular object tracking has been focused single-frame fiducial detection.
+A common assumption in these algorithms is that the image being processed is blur free, expectation which breaks down in empirical situations.
+For instance, the widely-used AprilTag [cite] marker and its successor, the AprilTag 2 [cite], rely on strong gradients between the tag's background and the surrounding white border to detect the tag.
+As shown in figure \ref{fig:blurred_tag}, oftentimes such algorithms fail to detect tags that have been blurred, either missing them entirely or discarding them as false positives as the identifying characteristics are completely blurred (for AprilTags this consists of a 6x6 binary grid in the center of the tag).
+
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=6cm, height=4cm]{blurred_tag}
+	\caption{A blurred AprilTag.
+Such tags are difficult for still frame based algorithms to detect as the payload (the ID encoded in the center of the tag) will cause the tag to be discarded as a false positive.}
+	\label{fig:blurred_tag}
+\end{figure}
+
+There has been considerable interest in designing blur resilient fiducials.
+Meghshyam et al.
+[cite] developed ringed markers to ensure that a region of the tag will always be perpendicular to the blur direction.
+However, this approach, while resistent to linear blur, does not take into account more complex blurs (e.g resulting from the tag itself pivoting around an axis).
+In addition, the increased resilience to blur comes at the cost of localization accuracy as only the center point of the tag can be reliably determined, whereas with other tags, such as the AprilTag, full pose estimation in 6 degrees of freedom (DOF) can be done based on just a single tag.
+
+Unlike previous attempts to engineer more resilient tags, we propose to exploit the temporal coherence between successive frames in a video sequence.
+As many high-blur scenarios, such as a quadroter-mounted moving camera or tracking an AprilTag-labeled moving object, involve a continuous stream of images taken at equidistant time intervals, we propose using prior knowledge on a fiducial's location to augment the detection algorithm.
+As such, this paper presents a novel particle-filter-based tracking algorithm (AprilTrack) capable of tracking AprilTags in 6 DOF.
+
+Previous work involving the tracking of blurred objects includes the Blur-driven Tracker (BLUT) framework proposed by Wu et al.
+[cite] for tracking motion-blurred targets.
+BLUT, a particle-filter based tracker, works on the observation that blur strength can be used estimate an object's motion and provide valuable information for successfully tracking a blurred object.
+While capable of successfully tracking blurred objects, BLUT suffers from the disadvantage that it can do so only in two dimensions and so cannot track a sequence of motion involving arbitrary rotations and translations.
+
+The novel method proposed in this paper is able to augment the AprilTag detector from Olson et al.
+[cite] with a particle-filter based tracker to successfully track an AprilTag through an arbitrary movement, providing valuable position information when existing methods would be unable to do so.
+In addition, the proposed algorithm is capable of incorporating the joint motion of an arbitrary number of tags to increase the accuracy of the tracked orientation.
+
+\textbf{Overview.} The remainder of the paper is organized as follows.
+Section 2 gives an overview of the method proposed in this paper and describes the algorithm in detail.
+Section 3 describes the implementation of the algorithm and evaluates its performance on a hand-labeled dataset of blurred moving tags with both individual tag motion and multi-tag ego motion.
+In sections 4 and 5 we discuss the limitations of this approach and suggest possible improvements for further research.
+
+\section{Method}
+
+The AprilTrack algorithm is similar in concept to the BLUT tracker by Wu et al.
+[cite] in that it uses a particle filter to integrate all available information about a marker's position.
+Overall, the description of the algorithm can be broken down into 3 parts: one describing the particle filter, another on the method for evaluating candidate particles, and another describing the modifications made for tracking in multi-tag scenarios.
+
+\subsection{The Particle Filter}
+
+The particle filter used for the AprilTrack algorithm is a Bayesian sequential importance resampling technique which approximates a posterior distribution of state variables.
+There are three major steps in the algorithm: prediction, update, and resampling.
+We use the vector  $\vec{x}_t$  to denote the 13 dimensional state vector describing the position, orientation, velocity, and rotational velocity of the tracked target at time $t$
+
+\begin{equation}
+\vec{x}_t = 
+\begin{bmatrix}
+	\vec{r} \\
+	\dot{\vec{r}} \\
+	\vec{q} \\
+	\vec{\omega}
+\end{bmatrix}
+\end{equation}
+
+where $\vec{r}$ is the three dimensional position of the tracked object, $\dot{\vec{q}}$ is the three dimensional velocity, $\vec{q}$ is a four dimensional unit quaternion containing the pose of the tracked object, and $\vec{\omega}$ is an angular velocity with components $\omega_x$, $\omega_y$, and $\omega_z$.
+We use $\vec{y}_k$ to denote the measurement (appearance) of the tag a particular time $t$.
+
+The predicting distribution can be formulated in a Bayesian setting as the computation of the distribution $p(\vec{x}_t|\vec{z}_{t-1})$ from the prior distribution $p(\vec{x}_{t-1}|\vec{z}_{1:t-1})$:
+
+\begin{equation} \label{eq:predict}
+p(\vec{x}_t|\vec{z}_{1:t-1}) = \int p(\vec{x}_t | \vec{x}_{t-1})p(\vec{x}_t|\vec{z}_{1:t-1})d\vec{x}_{t-1}
+\end{equation}
+
+Likewise, the update step can be formulated in a similar manner, where the prior distribution is updated with the latest measurement to obtain the posterior over $\vec{x}_t$:
+
+\begin{equation} \label{eq:update}
+p(\vec{x}_t|\vec{z}_{1:t}) \propto p(\vec{z}_t|\vec{x}_t)p(\vec{x}_t|\vec{z}_{1:t-1})
+\end{equation}
+
+In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$ using Monte-Carlo methods.
+A particle filter involves a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ where each sample is called a \emph{particle}.
+In the simplified case of a sequential importance resampling algorithm the prediction and update equations from \ref{eq:predict} and \ref{eq:update} can written as:
+
+\begin{equation}
+	\vec{x}^i_t \sim p(\vec{x}_t|\vec{x}^i_{t-1})
+\end{equation}
+
+\begin{equation}
+	w \propto p(\vec{z}_t|\vec{x}^i_t)
+\end{equation}
+
+After selecting the particle with the highest weight as the most likely candidate, the final resampling step simply consists of drawing a new set of $N$ particles with equal weights (normalized to one) from the distribution $p(\vec{x}_t|\vec{y}_{1:t})$ where $p(\vec{x}_t|\vec{y}_{1:t})$ is approximated by the set of weighted particles.
+
+\begin{equation}
+	p(\vec{x}_t|\vec{z}_{1:t}) \approx \sum_{i=1}^{N}{w^i_t\delta_{\vec{x}^i_t}}
+\end{equation}
+
+In AprilTrack, we model $p(\vec{x}_t|\vec{x}^i_{t-1})$ as a multivariate gaussian distribution with noise $\vec{\sigma}$ around 
+\begin{equation}
+\vec{\mu} = 
+\begin{bmatrix} 
+	\vec{r}^i_{t-1} + \Delta t * \dot{\vec{r}}^i_{t-1} \\
+ 	\dot{\vec{r}}^i_{t-1} \\
+	\vec{q}^i_{t-1} \dot{\vec{q}}^i_{t-1}\\
+	\vec{\omega}^i_{t-1}
+\end{bmatrix}
+\end{equation}
+
+where $\Delta t$ is the elapsed time and $\dot{\vec{q}}^i_{t-1}$ is the quaternion representation of $\Delta t * \vec{\omega}^i_{k-1}$ as described in [cite].
+We also use the \emph{noise quaternion} from [cite] to apply the gaussian noise to the quaternion $\vec{q}^i_{k-1} \dot{\vec{q}}^i_{k-1}$.
+For simplicity, we make the assumption that the input images have been taken at equidistant time intervals and incorporate $\Delta t$ into $\vec{\sigma}$.
+
+\subsection{Evaluation of Candidate Particles}
+
+The successful tracking of AprilTags in 6 DOF requires a method for evaluating the likelihood of candidate particles at arbitrary three dimensional positions and rotations.
+To this effect, the proposed method uses homographies to generate image \emph{patches}, which are unprojected representations of an AprilTag for a particular pose and image.
+
+Given the function $I(\begin{bmatrix} \lambda x, & \lambda y, & \lambda \end{bmatrix}^\intercal )$ representing the intensity of pixel at a particular $(x,y)$ of an image, a \emph{patch} in the image can be defined by the function
+
+\begin{equation}
+P(x_p,y_p| \vec{H}, I ) = I(\vec{H} * \begin{bmatrix} x_p & y_p & 1\end{bmatrix}^\intercal)
+\end{equation}
+
+where $x_p$ and $y_p$ are between $-1$ and $1$ and $\vec{H}$ is a 3x3 matrix (a homography) mapping from $(x_p, y_p, 1)$ to $(\lambda x, \lambda y, \lambda)$ which incorporates the pose of an AprilTag relative to the camera.
+
+This allows for the evaluation of candidate particles by the measuring the similarity between the particle's patch $P^i(x_p,y_p|\vec{H}^i,I)$ in the current image and the patch from the AprilTag's last successful detection, $P_{ref}(x_p,y_p|\vec{H}_{ref},I)$.
+Here, $H^i$ can be calculated with
+
+\begin{equation} \label{eq:particle_homography}
+\vec{H}^i = \begin{bmatrix}
+		\widehat{\vec{q}^i} & \vec{r}^i
+	\end{bmatrix} \begin{bmatrix}
+		\tau/2 & 0 & 0 \\
+		0 & \tau/2 & 0 \\
+		0 & 0 & 0 \\
+		0 & 0 & 1
+	\end{bmatrix}
+\end{equation}
+
+where $\widehat{\vec{q}}$ is the rotation matrix representation of particle's rotation quaternion, $\vec{r}$ is the particle's translation, and $\tau$ is the tag's edge length.
+The patch for the detected AprilTag uses the homography $\vec{H}_{ref}$ supplied by the AprilTag library, which computes the homography using the Direct Linear Transform algorithm [cite].
+
+To compare the patches efficiently, $P^i$ and $P_{ref}$ are evaluated at a regular grid from $(-1, -1)$ to $(1, 1)$ at intervals of $2 / \rho$.
+This results in two $\rho \times \rho$ matrices $\vec{M}^i$ and $\vec{M}_{ref}$.
+To compare the patches, the error between the two  matrices is calculated using the correlation coefficient between the individual elements of the matrices.
+
+\begin{equation}
+	\epsilon(\vec{A}, \vec{B}) = \frac{1}{2} - \frac{\sum_{i=1}^{\rho} \sum_{j=1}^{\rho} (\vec{A}_{ij} - \mu_{\vec{A}})(\vec{B}_{ij} - \mu_{\vec{B}}) }{2\sigma_{A} \sigma_{B}}
+\end{equation}
+
+where $\vec{A}$ and $\vec{B}$ are the matrix approximations of two patches, $\sigma_{\vec{A}}$ and $\sigma_{\vec{B}}$ are the standard deviations of the respective matrices' elements and $\mu_{\vec{A}}$ and $\mu_{\vec{B}}$ are their means.
+By normalizing by mean and standard deviation, using correlation as a similarity metric (as opposed to the sum of the absolute difference or squared difference) ensures lighting invariance for comparing patches containing the same marker under different illuminations.
+
+As in [cite] the patch error $\epsilon(\vec{M}^{i}, \vec{M}_{ref})$ is used to compute the observation likelihood $p(\vec{z}_t|\vec{x}_t)$ to update the weights and select the best particle with
+
+\begin{equation}
+	p(\vec{z}_t|\vec{x}^i_t) = exp(-\gamma \epsilon(\vec{M}^i, \vec{M}_{ref}))
+\end{equation}
+
+where the parameter $\gamma$ allows for more or less selective resampling.
+
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=7.5cm,height=5cm]{tracked_tag}
+	\caption{A blurred tag being tracked. The red border indicates that the tag was not detected with the AprilTag detector.}
+\end{figure}
+
+\begin{figure*}[t]
+	\label{fig:error}
+	\includegraphics[width=\textwidth,height=8cm]{Average_Error}
+	\caption{The average corner error between the hand-marked tag location and the tracked location.
+Much of the error from frames 0-200 comes from discrepancy between the tracker and the ground truth on the boundaries of blurred tags.
+At around frame 200, the tracker loses the tag due to a sudden rotation but is able to recover without needing a hard reset.}
+\end{figure*}
+
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=7.5cm,height=5cm]{Particles2}
+	\caption{The distribution of particles and their weights with respect to the X,Y position of the particles.}
+\end{figure}
+
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=7.5cm,height=5cm]{RefPatches}
+	\caption{The reference patches of various tags (by ID) in a multi-Tag tracking scenario.}
+\end{figure}
+
+\subsection{Multi-Tag Scenarios}
+
+To adapt this method to multi-tag scenarios, several extra parameters were introduced.
+In order to jointly track multiple tags, it is assumed that the relative poses of the tags are known.
+For each tag $T_k$ with ID $k$, the location $\vec{x}_{k}$ and orientation quaternion $\vec{g}_{k}$ are expressed such that
+
+\begin{equation}
+	\vec{X}_c = \widehat{\vec{q}^i} (\widehat{\vec{g}_k} \vec{X}_t + \vec{x}_k) + \vec{r}^i
+\end{equation}
+where $\vec{X}_t$ is a coordinate relative to the tag's reference frame, $\vec{X}_c$ is relative to the camera reference frame, and $\vec{q}^i$ and $\vec{r}^i$ are the position and orientation of particle $\vec{x}^i_t$.
+In the simple case of jointly moving tags, the $\vec{H}^i$ from equation \ref{eq:particle_homography} can then be rewritten as
+
+\begin{equation}
+	\vec{H}^i_{k} =
+	\begin{bmatrix}
+		\widehat{\vec{q}^i \vec{g_k}} & \widehat{\vec{q}} \vec{x_k} + \vec{r}^i
+	\end{bmatrix}\begin{bmatrix}
+		\tau_k/2 & 0 & 0 \\
+		0 & \tau_k/2 & 0 \\
+		0 & 0 & 0 \\
+		0 & 0 & 1
+	\end{bmatrix}
+\end{equation}
+
+where $\vec{H}^i_k$ represents the homography for particle $i$ and tag $k$ and $\tau_k$ is the edge length of tag $k$.
+
+In the ego-motion scenario, each particle's rotation and translation can be taken to be the camera's rotation and  translation such that
+
+\begin{equation}
+	\widehat{\vec{q}^i}\vec{X}_c + \vec{r}^i = \widehat{\vec{g}_k} \vec{X}_t + \vec{x}_k
+\end{equation}
+
+resulting in the homography $\vec{H}^i_k$ and the corresponding patch approximation $\vec{M}^i_k$ where
+
+\begin{equation}
+	\vec{H}^i_k = \begin{bmatrix}
+		\widehat{\vec{q}^i}^{-1} \widehat{\vec{g}_k} & \widehat{\vec{q}^i}^{-1} (\vec{x}_k - \vec{r}^i)
+	\end{bmatrix} \begin{bmatrix}
+		\tau_k/2 & 0 & 0 \\
+		0 & \tau_k/2 & 0 \\
+		0 & 0 & 0 \\
+		0 & 0 & 1
+	\end{bmatrix}
+\end{equation}
+
+In both ego-motion and joint tag motion, the observation likelihood $p(\vec{z}_t|\vec{x}_{t})$ then becomes
+
+\begin{equation}
+	p(\vec{z}_t|\vec{x}^i_{t}) = exp(\sum^{K}_{k=1} -\gamma \epsilon(\vec{M}^i_k, \vec{M}_{k, ref}))
+\end{equation}
+
+where $K$ is the number of tags being tracked and $\vec{M}_{k,ref}$ is the reference patch approximation for tag $k$ from the last successful detection.
+
+\section{Results and Discussion}
+
+The AprilTrack algorithm was primarily evluated on a manually-labeled single-tag sequence with a moving tag.
+The algorithm was also run on an ego-motion multi-tag sequence which will be qualitatively discussed.
+
+\footnote{Videos of the sequences and the tracker's performance are available online at https://www.youtube.com/watch?v=BmZLKFBkcxU, https://www.youtube.com/watch?v=ebWaxO7JhKE, and }
+
+The specific implementation of the AprilTrack algorithm used in this experiment was written in MATLAB.
+For both scenarios, the parameters $\gamma = 10$, $\rho=32$ were used.
+$\tau$ was $0.1936$ m for all tags, slightly larger than the actual tag size of $0.1635$ m to include some of the whitespace around the tag.
+In the single-tag scenario, the standard deviation $\vec{\sigma}$  of the gaussian noise used in the state transition distribution $p(\vec{x}_t|\vec{x}^i_{t-1})$ was $0.01$ for the $\vec{r}$ component, $0.05$ for $\vec{q}$, $0.02$ for $\dot{\vec{r}}$, and $0$ for $\omega$.
+
+When evaluated on the single-tag sequence, the algorithm ran with $N=6000$ at around $0.5$ fps on an i7 Haswell Intel CPU.
+In experiments with fewer particles, it was found that the algorithm could track a tag well with as few as $1000$ particles with around $5$ fps.
+With a more optimized C++ implementation, the framerate could likely be brought up to $10-15$ fps, making the algorithm (run with less agressive parameters) a possible candidate for real time scenarios.
+Currently, the MATLAB implementation is intended for offline post-processing of a sequence.
+
+The pixel corner error hovers at around 10 pixels for most of the sequence.
+The tracker loses the tag momentarily during a quick twist at around frame 200 but is able to recover rapidly without any manual intervention.
+Much of the error observed throughout the sequence is the result of disagreement on the boundary of a blurred tag rather than a systematic issue with the tracker itself.
+
+When running with a multi-tag configuration, the tracker is able to make use of the additional information to more accurately track the camera's location than in a single-tag scenario.
+The main challenge for running the algorithm with multiple tags was not successfully tracking the tags---the algorithm was able to do that with half the particles of the single-tag scenario.
+Rather, the majority of the reprojection error arises from error in the relative tag positions.
+
+\section{Conclusion}
+
+This paper has presented the AprilTrack algorithm augmenting the standard AprilTag detection algorithm with a novel tracker capable of tracking the tags in 6 DOF by using homographies to evaluate candidate particles.
+Experimental results show that the tracker is capable of accurately tracking both single tags and consolidating information from multiple tags in ego-motion and joint tag-motion scenarios.
+In future research we plan to investigate other possible error metrics for patch comparison and look to deploy AprilTrack in realtime scenarios.
 

### 2017-01-02 AprilTrack.m

@@ -20,9 +20,8 @@ classdef AprilTrack < TagSource
             
             for i=1:tagParams.patchSize(2)
                 for j=1:tagParams.patchSize(1)
-                    idx = [i / tagParams.patchSize(2) - 0.5; ...
-                           j / tagParams.patchSize(1) - 0.5];
-
+                    idx = 2* [i / tagParams.patchSize(2) - 0.5; ...
+                              j / tagParams.patchSize(1) - 0.5];
                     coordinates(j, i, :) = idx;
                 end
             end
@@ -53,9 +52,9 @@ classdef AprilTrack < TagSource
                 tags = [tags mm.process(img, tagMap)];
             end
         end
-        function debug(this, fig1, fig2, fig3, fig4)
+        function debug(this, img, fig1, fig2, fig3, fig4, fig5)
             if length(this.motionModels) > 0
-                this.motionModels{1}.debug(fig1, fig2, fig3, fig4);
+                this.motionModels{1}.debug(img, fig1, fig2, fig3, fig4, fig5);
             end
         end
     end

### 2017-01-02 MotionModel.m

@@ -124,7 +124,8 @@ classdef MotionModel < handle
 
                     H = homography_solve(pin, tag.corners);
                     % Scale to get the border as well + right units
-                    s = this.modelledTags{i}.size + this.modelledTags{i}.border;
+                    s = (this.modelledTags{i}.size + this.modelledTags{i}.border) / 2;
+                    
                     S = [s(1) 0 0; 0 s(2) 0; 0 0 1];
                     
                     tag.homography = H * S;
@@ -325,7 +326,7 @@ classdef MotionModel < handle
         end
         
         
-        function debug(this, fig1, fig2, fig3, fig4)
+        function debug(this, img, fig1, fig2, fig3, fig4, fig5)
             sfigure(fig1);
             colormap(gray(255));
             
@@ -374,6 +375,29 @@ classdef MotionModel < handle
             sfigure(fig4);
             scatter3(gen_particles(1, :), gen_particles(2, :), ...
                      gen_m, 5, gen_m);
+                 
+            sfigure(fig5);
+            colormap(gray(255));
+
+            n = round(sqrt(length(this.modelledTags))) + 1;
+            for i=1:length(this.modelledTags)
+                t = this.modelledTags{i};
+                
+                w = this.visibleTags(i);
+                if w > 0
+                    % Transform the state
+                    t.state = this.transform(t.state, this.x);
+                    t.homography = homography_from_state(this.tagParams.K, t);
+                    
+                    p = extract_patch(this.tagParams.patchSize, ...
+                                      this.tagParams.coords, ...
+                                      img, t);
+                    
+                    h = subplot(n, n, i);
+                    image(p, 'Parent', h);
+                    title(t.id);
+                end
+            end
         end        
     end
 end

### 2017-01-02 algorithm.m

@@ -1,5 +1,5 @@
 function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, record)
-    imgs = [];
+    imgs = {};
     corners = [];
     pos = [];
     rot = [];
@@ -11,6 +11,7 @@ function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, reco
     fig3 = sfigure(3);
     fig4 = sfigure(4);
     fig5 = sfigure(5);
+    fig6 = sfigure(6);
     
     figure(1);
     
@@ -23,6 +24,9 @@ function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, reco
         counter = counter + 1;
         if mod(counter, 10) == 0
             clc;
+        end
+        if (counter == 1000)
+            return;
         end
         
         fprintf('------------------------------\n');      
@@ -43,6 +47,7 @@ function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, reco
         clf(fig3);
         clf(fig4);
         clf(fig5);
+        clf(fig6);
         
         fprintf('// Took %f\n', toc());
         fprintf(':: Displaying result\n');
@@ -70,7 +75,7 @@ function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, reco
         fprintf(':: Displaying debug stuff\n');
         tic();
 
-        tracker.debug(fig2, fig3, fig4, fig5);
+        tracker.debug(img, fig2, fig3, fig4, fig5, fig6);
         sfigure(1);
 
         fprintf('// Took %f\n', toc());
@@ -80,7 +85,7 @@ function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, reco
             tic();
             fprintf('Getting frame');
 
-            imgs = [ imgs getframe(fig1) ];
+            imgs{counter} = getframe(fig1);
 
             disp('-- '), disp(toc());
         end

### 2017-01-02 config/movcam1_config.m

@@ -18,7 +18,7 @@ function [tracker, detector, frames] = movcam1_config(skip_rate)
     tracker = AprilTrack(trackParams);
 
     % Add a motion model
-    mmParams(1).err_discard_threshold = 1;
+    mmParams(1).err_discard_threshold = 1.1;
     mmParams(1).num_particles = 3000;
     mmParams(1).process_noise = [0.01 0.01 0.01 ...
                                  0.01 0.01 0.01 ...
@@ -33,8 +33,8 @@ function [tracker, detector, frames] = movcam1_config(skip_rate)
 
     model.loadTags('../data/movcam_tags.txt');
     
-    initial = [-0.424; -0.7353; 3.4589; ...
-               0.9791; 0.1296; -0.1559; -0.0135; ...
+    initial = [-0.6899; -0.1881; -3.4893; ...
+               0.9791; -0.1296; 0.1559; 0.0135; ...
                0; 0; 0;  0; 0; 0];
     
     model.initializeParticlesTo(initial);

### 2017-01-02 config/test_video_config.m

@@ -21,8 +21,8 @@ function [tracker, detector, frames] = test_video_config()
     mmParams(1).num_particles = 6000;
     mmParams(1).process_noise = [0.01 0.01 0.01 ...
                                  0.05 0.05 0.05 ...
-                                 0.02 0.02 0.02 ...
-                                 0.00 0.00 0.00];
+                                 0.01 0.01 0.01 ...
+                                 0.01 0.01 0.01];
 
     % noise = 1 / (k + alpha * w) * process_noise * randn
     % Allows for particles of higher/lower weight to have
@@ -34,7 +34,7 @@ function [tracker, detector, frames] = test_video_config()
 
     % For measurement error --> weight conversion
     % where weight = e^(-lambda * measurement)
-    mmParams(1).lambda = 9;
+    mmParams(1).lambda = 10;
 
     model = MotionModel(mmParams, @transform_tag);
 

### 2017-01-02 run.m

@@ -1,8 +1,8 @@
-function [ imgs, corners, pos, rot ] = run(config, gt, shouldRecord)
+function [ imgs, corners, pos, rot ] = run(config, shouldRecord)
     tracker = [];
     detector = [];
     frames = [];
-    imgs = [];
+    imgs = {};
     corners = [];
     pos = [];
     rot = [];
@@ -24,8 +24,8 @@ function [ imgs, corners, pos, rot ] = run(config, gt, shouldRecord)
     end
     
     if (exist('shouldRecord', 'var') && shouldRecord)
-        [imgs, corners, pos, rot] = algorithm(tracker, detector, frames, gt, true);
+        [imgs, corners, pos, rot] = algorithm(tracker, detector, frames, true);
     else
-        algorithm(tracker, detector, frames, false);
+        [imgs, corners, pos, rot] = algorithm(tracker, detector, frames, false);
     end
 end

### 2017-01-02 utils/homography_from_state.m

@@ -6,7 +6,7 @@ function [ H ] = homography_from_state(K, tag)
     % Create the homography
     
     % Project the coords
-    s = tag.size + tag.border;
+    s = (tag.size + tag.border) / 2;
     S = [s(1) 0 0; ...
          0 s(2) 0; ...
          0 0    1];

### 2017-01-02 utils/transform_ego.m

@@ -1,9 +1,9 @@
 function newPose = transform_ego(tagPose, x)
     % Local -> World transform
     WL = [[quat_to_rotm(tagPose(4:7)') [tagPose(1); tagPose(2); tagPose(3)]]; [0 0 0 1]];
-    % World -> Camera transform
-    CW = [[quat_to_rotm(x(4:7)') [x(1); x(2); x(3)]]; [0 0 0 1]];
-    T = CW * WL;
+    % Camera -> World transform
+    WC = [[quat_to_rotm(x(4:7)') [x(1); x(2); x(3)]]; [0 0 0 1]];
+    T = inv(WC) * WL;
     newPose = tagPose;
     newPose(1:3) = [T(1, 4) T(2, 4) T(3, 4)];
     newPose(4:7) = rotm_to_quat(T(1:3, 1:3));

### 2017-01-09 paper/paper.tex
Working on paper

@@
 
-Visual tracking of fiducial markers (or fiducials) plays an important role in many robotics and computer vision related fields, including augmented and virtual reality, visual odometry, and object tracking.
-Due to these scenarios taking place in tructured indoors environments where illumination is often limited, long exposure times to compensate for the reduced lighting produce blurred images, especially when taken from a mobile platform such as quadrotor.
+Visual tracking of fiducial markers (or simply fiducials) plays an important role in many robotics and computer vision related fields and applications, including augmented and virtual reality, visual odometry, and object tracking in 6 degrees of freedom (DOF).
+However, as many of these applications involve the of fiducials in structured indoors environments where illumination is often limited, long exposure times to compensate for the reduced lighting produce blurred images, especially when taken from a mobile platform such as quadrotor.
 
- The majority of previous literature in fiducial detection and monocular object tracking has been focused single-frame fiducial detection.
+ The majority of previous literature in fiducial detection has been focused single-frame fiducial detection.
 A common assumption in these algorithms is that the image being processed is blur free, expectation which breaks down in empirical situations.

### 2017-01-14 paper/paper.tex
Worked on paper a bit

--- 2017-01-02 paper/paper.tex
@@
 	The AprilTrack algorithm proposed in this paper seeks to bridge two existing fields of research in computer vision literature: fiducial detection and monocular object tracking.
-With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth position data for mobile robots, there has been a considerable interest in designing robust and blur-resistant fiducials.
-Rather than devising a new fiducial, This paper attempts to improve on existing fiducials such as the popular AprilTag marker by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom, allowing for tag-based localization in situations where still frame based methods would be unable to do so.
-This novel particle-filter based algorithm is evaluated on several datasets involving both ego-motion and tag-based-motion and is compared to standalone tag detection algorithms and ground-truth tag position data.
+With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth position data for mobile robots, there has been considerable interest in designing robust and blur-resistant fiducials.
+Rather than devising a new fiducial, this paper attempts to improve on existing fiducials such as the popular AprilTag marker by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom, allowing for tag-based localization in situations where still frame based methods would be unable to do so.
+This novel particle-filter based algorithm is evaluated on several datasets involving both ego-motion and tag-based-motion and is compared to ground-truth tag position data.
 
@@
 where $\vec{r}$ is the three dimensional position of the tracked object, $\dot{\vec{q}}$ is the three dimensional velocity, $\vec{q}$ is a four dimensional unit quaternion containing the pose of the tracked object, and $\vec{\omega}$ is an angular velocity with components $\omega_x$, $\omega_y$, and $\omega_z$.
-We use $\vec{y}_k$ to denote the measurement (appearance) of the tag a particular time $t$.
+We use $\vec{z}_k$ to denote the measurement (appearance) of the tag a particular time $t$.
 
@@
 
-After selecting the particle with the highest weight as the most likely candidate, the final resampling step simply consists of drawing a new set of $N$ particles with equal weights (normalized to one) from the distribution $p(\vec{x}_t|\vec{y}_{1:t})$ where $p(\vec{x}_t|\vec{y}_{1:t})$ is approximated by the set of weighted particles.
+After selecting the particle with the highest weight as the most likely candidate, the final resampling step simply consists of drawing a new set of $N$ particles with equal weights (normalized to one) from the distribution $p(\vec{x}_t|\vec{z}_{1:t})$ where $p(\vec{x}_t|\vec{z}_{1:t})$ is approximated by the set of weighted particles.
 
@@
 
-As in [cite] the patch error $\epsilon(\vec{M}^{i}, \vec{M}_{ref})$ is used to compute the observation likelihood $p(\vec{z}_t|\vec{x}_t)$ to update the weights and select the best particle with
-
-\begin{equation}
-	p(\vec{z}_t|\vec{x}^i_t) = exp(-\gamma \epsilon(\vec{M}^i, \vec{M}_{ref}))
-\end{equation}
-
-where the parameter $\gamma$ allows for more or less selective resampling.
-
 \begin{figure}[b]
@@
 	\centering
-	\includegraphics[width=7.5cm,height=5cm]{Particles2}
+	\includegraphics[width=7cm,height=4.75cm]{Particles2}
 	\caption{The distribution of particles and their weights with respect to the X,Y position of the particles.}
@@
 	\centering
-	\includegraphics[width=7.5cm,height=5cm]{RefPatches}
+	\includegraphics[width=7cm,height=4.75cm]{RefPatches}
 	\caption{The reference patches of various tags (by ID) in a multi-Tag tracking scenario.}
 \end{figure}
+
+As in [cite] the patch error $\epsilon(\vec{M}^{i}, \vec{M}_{ref})$ is used to compute the observation likelihood $p(\vec{z}_t|\vec{x}_t)$ to update the weights and select the best particle with
+
+\begin{equation}
+	p(\vec{z}_t|\vec{x}^i_t) = exp(-\gamma \epsilon(\vec{M}^i, \vec{M}_{ref}))
+\end{equation}
+
+where the parameter $\gamma$ allows for more or less selective resampling.
 

### 2017-01-14 MotionModel.m

@@ -121,7 +121,6 @@ classdef MotionModel < handle
                             tagSize(1), -tagSize(2); ...
                             tagSize(1),  tagSize(2);...
                            -tagSize(1),  tagSize(2)];                
-
                     H = homography_solve(pin, tag.corners);
                     % Scale to get the border as well + right units
                     s = (this.modelledTags{i}.size + this.modelledTags{i}.border) / 2;
@@ -217,7 +216,7 @@ classdef MotionModel < handle
             %%{
             [z, i] = max(this.weights);
             this.x = this.particles(:, i);
-            
+            this.x
             % Measure which tags we should get rid of
             this.removeOccludedTags(this.x, img);
 
@@ -314,10 +313,11 @@ classdef MotionModel < handle
                                       this.tagParams.coords, ...
                                       img, t);
                     
-                    %p
-                    %t.refPatch
                     err = measure_patch_error(p, t.refPatch);
-                    err
+                    
+                    % For error debugging
+                    % err
+                    
                     if err >= this.params.err_discard_threshold
                         this.visibleTags(i) = 0;
                     end

### 2017-01-14 config/movcam1_config.m

@@ -19,11 +19,15 @@ function [tracker, detector, frames] = movcam1_config(skip_rate)
 
     % Add a motion model
     mmParams(1).err_discard_threshold = 1.1;
-    mmParams(1).num_particles = 3000;
-    mmParams(1).process_noise = [0.01 0.01 0.01 ...
-                                 0.01 0.01 0.01 ...
-                                 0.01 0.01 0.01 ...
-                                 0.01 0.01 0.01];
+    mmParams(1).num_particles = 5000;
+    mmParams(1).process_noise = [0.0 0.0 0.0 ...
+                                 0.1 0.1 0.1 ...
+                                 0.0 0.0 0.0 ...
+                                 0.0 0.0 0.0];
+%    mmParams(1).process_noise = [0.02 0.02 0.02 ...
+%                                 0.02 0.02 0.02 ...
+%                                 0.01 0.01 0.01 ...
+%                                 0.01 0.01 0.01];
 
     % For measurement error --> weight conversion
     % where weight = e^(-lambda * measurement)
@@ -33,8 +37,8 @@ function [tracker, detector, frames] = movcam1_config(skip_rate)
 
     model.loadTags('../data/movcam_tags.txt');
     
-    initial = [-0.6899; -0.1881; -3.4893; ...
-               0.9791; -0.1296; 0.1559; 0.0135; ...
+    initial = [-0.6899; -0.1881; 3.4893; ...
+               0.1540;  -0.0023; 0.9792; 0.1319; ...
                0; 0; 0;  0; 0; 0];
     
     model.initializeParticlesTo(initial);

### 2017-01-14 save_images.m

@@ -1,6 +1,6 @@
 function save_images(input, output_dir)
     for c=1:length(input)
-        img = input(c).cdata;
+        img = input{c}.cdata;
         file = sprintf('%s/%04d.png', output_dir, c);
         imwrite(img, file);
     end

### 2017-01-14 utils/homography_from_state.m

@@ -3,6 +3,7 @@ function [ H ] = homography_from_state(K, tag)
     R = quat_to_rotm(tag.state(4:7));
     T = tag.state(1:3);
     H = K * [R(:, 1:2) T];
+    %H = [R(:, 1:2) T];
     % Create the homography
     
     % Project the coords

### 2017-01-14 utils/homography_project.m

@@ -1,6 +1,11 @@
 
 function [ x ] = homography_project(H, X)
     t = H * X;
+    
+    if any(t(3,:) <= 0)
+        x = ones([2 size(t,2)]) * -1;
+    end
+    
     % Divide by the last row
     x_x = t(1, :) ./ t(3, :);
     x_y = t(2, :) ./ t(3, :);

### 2017-01-17 paper/paper.tex
Paper in submittable state

--- 2017-01-14 paper/paper.tex
@@
 
+\usepackage{cite}
 \usepackage{graphicx}
@@
 \renewcommand{\vec}[1]{\boldsymbol{#1}}
+
+\renewcommand{\eqref}[1]{\textup{{\normalfont(\ref{#1}}\normalfont)}}
 
@@
 	The AprilTrack algorithm proposed in this paper seeks to bridge two existing fields of research in computer vision literature: fiducial detection and monocular object tracking.
-With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth position data for mobile robots, there has been considerable interest in designing robust and blur-resistant fiducials.
-Rather than devising a new fiducial, this paper attempts to improve on existing fiducials such as the popular AprilTag marker by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom, allowing for tag-based localization in situations where still frame based methods would be unable to do so.
-This novel particle-filter based algorithm is evaluated on several datasets involving both ego-motion and tag-based-motion and is compared to ground-truth tag position data.
+With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth pose estimation for visual odometry datasets and mobile robots, there has been considerable interest in designing robust and blur-resistant fiducials.
+Rather than devising a new fiducial, this work attempts to improve on existing fiducials, such as the popular AprilTag marker, by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom, allowing for tag-based localization in situations where current detectors would be unable to do so.
+This novel patch-based particle filter tracking algorithm, which is extended to simulatenously track multiple tags in ego-motion scenarios, is qualitatively and quantitatively evaluated on several AprilTag sequences, demonstrating that it is capable of reliably localizing blurred tags.
 
@@
 
-Visual tracking of fiducial markers (or fiducials) plays an important role in many robotics and computer vision related fields, including augmented and virtual reality, visual odometry, and object tracking.
-Due to these scenarios taking place in tructured indoors environments where illumination is often limited, long exposure times to compensate for the reduced lighting produce blurred images, especially when taken from a mobile platform such as quadrotor.
+Visual tracking of fiducial markers (or simply fiducials) plays an important role in many robotics and computer vision related fields and applications, including augmented and virtual reality \cite{AugmentedReality}, visual odometry/SLAM \cite{TagSLAM}, and object tracking.
+However, as many of these applications involve the use of fiducials in structured indoors environments where illumination is often limited, long exposure times to compensate for the reduced lighting produce blurred images, especially when taken from a mobile platform such as a quadrotor or a handheld device.
 
- The majority of previous literature in fiducial detection and monocular object tracking has been focused single-frame fiducial detection.
-A common assumption in these algorithms is that the image being processed is blur free, expectation which breaks down in empirical situations.
-For instance, the widely-used AprilTag [cite] marker and its successor, the AprilTag 2 [cite], rely on strong gradients between the tag's background and the surrounding white border to detect the tag.
-As shown in figure \ref{fig:blurred_tag}, oftentimes such algorithms fail to detect tags that have been blurred, either missing them entirely or discarding them as false positives as the identifying characteristics are completely blurred (for AprilTags this consists of a 6x6 binary grid in the center of the tag).
+The focus of this work is primarily on providing better pose estimation for visual odometry datasets which use AprilTags placed to known locations to provide the ground truth pose data.
+As the pose estimation in this scenario can be done in a post-processing step after the data has been collected, the current implementation of AprilTrack is not designed to run in realtime.
+We intend to explore the use of AprilTrack in realtime scenarios in future research.
+
+\subsection{Previous Work}
+
+ The majority of previous literature in fiducial detection has been focused single-frame fiducial detection \cite{RUNETag, BlurResilient, AprilTag}.
+A common assumption in these algorithms is that the image being processed shows little motion blur, an expectation which breaks down in empirical situations.
+For instance, the widely-used AprilTag marker \cite{AprilTag} and its successor, the AprilTag 2 \cite{AprilTag_2}, rely on strong gradients between the tag's background and the surrounding white border to detect the tag.
+As shown in figure \ref{fig:blurred_tag}, oftentimes such algorithms fail to detect tags that have been blurred, either missing them entirely or mistakenly discarding them when eliminating false positives through some encoded fiducial ID (for AprilTags this consists of a 6x6 binary grid in the center of the tag which is used to verify the validity of the detection).
 
@@
 
-There has been considerable interest in designing blur resilient fiducials.
+Previously, there has been interest in designing blur resilient fiducials.
 Meghshyam et al.
-[cite] developed ringed markers to ensure that a region of the tag will always be perpendicular to the blur direction.
-However, this approach, while resistent to linear blur, does not take into account more complex blurs (e.g resulting from the tag itself pivoting around an axis).
-In addition, the increased resilience to blur comes at the cost of localization accuracy as only the center point of the tag can be reliably determined, whereas with other tags, such as the AprilTag, full pose estimation in 6 degrees of freedom (DOF) can be done based on just a single tag.
+\cite{BlurResilient} have developed ringed markers to ensure that a region of the tag will always be perpendicular to the blur direction.
+However, this approach, while resistent to linear blur, does not take into account more complex blurs (e.g resulting from the tag pivoting around the vertical axis).
+In addition, the increased resilience to blur comes at the cost of localization accuracy as only the center point of the ringed tag can be reliably determined \cite{BlurResilient}, whereas with other fiducials, such as the AprilTag \cite{AprilTag}, full pose estimation in 6 degrees of freedom (DOF) can be done based on just a single tag.
+Another significant disadvantage of the ringed marker from Meghshyam et al.
+is the limited number of tag ID's, as their proposed encoding method allows for a maximum of 4 distinct tags \cite{BlurResilient}.
 
-Unlike previous attempts to engineer more resilient tags, we propose to exploit the temporal coherence between successive frames in a video sequence.
-As many high-blur scenarios, such as a quadroter-mounted moving camera or tracking an AprilTag-labeled moving object, involve a continuous stream of images taken at equidistant time intervals, we propose using prior knowledge on a fiducial's location to augment the detection algorithm.
-As such, this paper presents a novel particle-filter-based tracking algorithm (AprilTrack) capable of tracking AprilTags in 6 DOF.
+Unlike previous attempts to engineer more resilient tags, the algorithm proposed here exploits the temporal coherence between successive frames in a video sequence to track a tag's location between successful detections.
+As many high-blur scenarios involve a continuous stream of images taken at equidistant time intervals (e.g a quadroter-mounted moving camera viewing tags or a stationary camera tracking an AprilTag-labeled moving object), we propose to use knowledge of a fiducial's prior location to augment the detection algorithm.
+As such, this paper presents a novel particle-filter-based tracking algorithm (AprilTrack) capable of tracking AprilTags in 6 DOF with both joint-tag motion and camera ego-motion.
 
 Previous work involving the tracking of blurred objects includes the Blur-driven Tracker (BLUT) framework proposed by Wu et al.
-[cite] for tracking motion-blurred targets.
-BLUT, a particle-filter based tracker, works on the observation that blur strength can be used estimate an object's motion and provide valuable information for successfully tracking a blurred object.
-While capable of successfully tracking blurred objects, BLUT suffers from the disadvantage that it can do so only in two dimensions and so cannot track a sequence of motion involving arbitrary rotations and translations.
+\cite{BLUT} for tracking motion-blurred targets.
+BLUT, a particle-filter based tracker, works on the observation that blur strength and direction can be used estimate an object's velocity and provide valuable information for successfully tracking a blurred object.
+While capable of reliably tracking blurred objects, BLUT suffers from the disadvantage that it can do so only in two dimensions (i.e it is assumed that the object always faces the camera) and so cannot track a sequence of involving arbitrary tag rotations and translations.
 
-The novel method proposed in this paper is able to augment the AprilTag detector from Olson et al.
-[cite] with a particle-filter based tracker to successfully track an AprilTag through an arbitrary movement, providing valuable position information when existing methods would be unable to do so.
-In addition, the proposed algorithm is capable of incorporating the joint motion of an arbitrary number of tags to increase the accuracy of the tracked orientation.
+The novel method proposed in this paper augments the AprilTag detection algorithm originally designed by Olson \cite{AprilTag} and improved by Wang et al \cite{AprilTag_2} with a particle-filter based algorithm to reliably track an AprilTag through an arbitrary sequence of images, providing valuable fiducial-based position information when existing methods would be unable to do so.
+In addition, the proposed algorithm is extended to incorporate the joint motion of an arbitrary number of tags to increase the accuracy of the tracked orientation and position.
 
@@
 Section 2 gives an overview of the method proposed in this paper and describes the algorithm in detail.
-Section 3 describes the implementation of the algorithm and evaluates its performance on a hand-labeled dataset of blurred moving tags with both individual tag motion and multi-tag ego motion.
-In sections 4 and 5 we discuss the limitations of this approach and suggest possible improvements for further research.
+Section 3 describes the implementation of the algorithm and evaluates its performance on a manually-labeled dataset of blurred moving tags both quantitatively and qualitatively.
+In section 4 we discuss the limitations of this approach and suggest possible avenues of future research.
 
@@
 The AprilTrack algorithm is similar in concept to the BLUT tracker by Wu et al.
-[cite] in that it uses a particle filter to integrate all available information about a marker's position.
-Overall, the description of the algorithm can be broken down into 3 parts: one describing the particle filter, another on the method for evaluating candidate particles, and another describing the modifications made for tracking in multi-tag scenarios.
+\cite{BLUT} in that it uses a particle filter to integrate all available information about a marker's position.
+Overall, the description of the algorithm can be broken down into three parts: one describing the particle filter, another on the method for evaluating candidate particles, and one describing the modifications made to simulatenously track multiple tags.
 
@@
 
-The particle filter used for the AprilTrack algorithm is a Bayesian sequential importance resampling technique which approximates a posterior distribution of state variables.
+The particle filter used for the AprilTrack algorithm is a Bayesian Sequential Importance Resampling (SIR) technique which approximates a posterior distribution of state variables at discrete time intervals.
 There are three major steps in the algorithm: prediction, update, and resampling.
@@
 
-where $\vec{r}$ is the three dimensional position of the tracked object, $\dot{\vec{q}}$ is the three dimensional velocity, $\vec{q}$ is a four dimensional unit quaternion containing the pose of the tracked object, and $\vec{\omega}$ is an angular velocity with components $\omega_x$, $\omega_y$, and $\omega_z$.
-We use $\vec{z}_k$ to denote the measurement (appearance) of the tag a particular time $t$.
+where $\vec{r}$ is the three dimensional position of the tracked object, $\dot{\vec{r}}$ is the three dimensional velocity, $\vec{q}$ is a four dimensional unit quaternion containing the pose of the tracked object, and $\vec{\omega}$ is an angular velocity with components $\omega_x$, $\omega_y$, and $\omega_z$.
 
-The predicting distribution can be formulated in a Bayesian setting as the computation of the distribution $p(\vec{x}_t|\vec{z}_{t-1})$ from the prior distribution $p(\vec{x}_{t-1}|\vec{z}_{1:t-1})$:
+\textbf{Optimal filtering.} The prediction step can be formulated in a Bayesian setting as the computation of the prior distribution $p(\vec{x}_{t}|\vec{z}_{t-1})$ from the filtering distribution $p(\vec{x}_{t-1}|\vec{z}_{1:t-1})$ where $\vec{z}_t$ represents the measurement (appearance of a fiducial) taken at time $t$.
+As in \cite{ParticleNotes}:
 
 \begin{equation} \label{eq:predict}
-p(\vec{x}_t|\vec{z}_{1:t-1}) = \int p(\vec{x}_t | \vec{x}_{t-1})p(\vec{x}_t|\vec{z}_{1:t-1})d\vec{x}_{t-1}
+p(\vec{x}_t|\vec{z}_{1:t-1}) = \int p(\vec{x}_t | \vec{x}_{t-1})p(\vec{x}_{t-1}|\vec{z}_{1:t-1})d\vec{x}_{t-1}
 \end{equation}
 
-Likewise, the update step can be formulated in a similar manner, where the prior distribution is updated with the latest measurement to obtain the posterior over $\vec{x}_t$:
-
+Likewise, the update step can be formulated in a similar manner, where the prior distribution calculated in the prediction step is updated with the latest measurement to obtain the posterior over $\vec{x}_t$.
+Mathematically this can be derived using Bayes' law to express $p(\vec{x}_t|\vec{z}_{1:t})$ in terms of $p(\vec{z}_t|\vec{x}_t)$ and $p(\vec{x}_t|\vec{z}_{1:t-1})$.
+Using Bayes' law yields
+\begin{equation*}
+	p(\vec{x}_t|\vec{z}_{1:t}) = p(\vec{x}_t|\vec{z}_t,\vec{z}_{1:t-1})
+\end{equation*}
+\begin{equation*}
+p(\vec{x}_t|\vec{z}_t,\vec{z}_{1:t-1}) = \frac{p(\vec{z}_t|\vec{x}_t, \vec{z}_{1:t-1})p(\vec{x}_t|\vec{z}_{1:t-1})p(\vec{z}_{1:t-1})}{p(\vec{z}_{1:t})}
+\end{equation*}
+As the model $p(\vec{x}_t|\vec{x}_{t-1})$ is Markovian, the current measurement is conditionally independent of previous measurements given $\vec{x}_t$, meaning that $p(\vec{z}_{t}|\vec{x}_t, \vec{z}_{1:t-1})$ can be written as $p(\vec{z}_t|\vec{x}_t)$.
+$p(\vec{z}_{1:t})$ can also be expressed as $p(\vec{z}_t|\vec{z}_{1:t-1})p(\vec{z}_{1:t-1})$ resulting in the final update equation \cite{ParticleNotes}
 \begin{equation} \label{eq:update}
-p(\vec{x}_t|\vec{z}_{1:t}) \propto p(\vec{z}_t|\vec{x}_t)p(\vec{x}_t|\vec{z}_{1:t-1})
+	p(\vec{x}_t|\vec{z}_{1:t}) = \frac{p(\vec{z}_{t}|\vec{x}_t)p(\vec{x}_{t}|\vec{z}_{1:t-1})}{p(\vec{z}_t|\vec{z}_{1:t-1})}
 \end{equation}
 
-In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$ using Monte-Carlo methods.
-A particle filter involves a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ where each sample is called a \emph{particle}.
-In the simplified case of a sequential importance resampling algorithm the prediction and update equations from \ref{eq:predict} and \ref{eq:update} can written as:
+\textbf{Particle Filters.} In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods.
+A particle filter consists of a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ drawn from the proposal distribution $q(\vec{x}^i)$ where each sample is called a \emph{particle}.
+This allows for the filtering distribution $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$ to be approximated by (as described in \cite{ParticleNotes})
+
+\begin{equation}
+ 	p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1}) \approx \sum_{i=1}^{N}{w^i_{t-1}\delta_{\vec{x}^i_{0:t-1}}}
+\end{equation}
+
+ where
+\begin{equation}
+	\sum_{i=1}^{N}{w^i_t} = 1
+\end{equation}
+
+As the particles are drawn from the proposal distribution $q(\vec{x})$, the sample weight $w^i$ must take into account any discrepancy between $p(\vec{x})$ and $q(\vec{x})$ so that $w^i_t \propto p(\vec{x}^i_{1:t}|\vec{z}_{1:t})/q(\vec{x}^i_{1:t}|\vec{z}_{1:t})$.
+
+In the simplified case of a Sequential Importance Resampling algorithm, the prediction and update equations derived from \eqref{eq:predict} and \eqref{eq:update} can, in the context of a SIR particle filter, be simplified to:
 
@@
 \end{equation}
+where $\vec{x}^i_t \sim p(\vec{x}_t|\vec{x}^i_{t-1})$ denotes that, during the prediction step, the particles for the prior distribution at time $t$ are sampled from the known distribution $p(\vec{x}_t|\vec{x}^i_{t-1})$.
+The distribution $p(\vec{z}_t|\vec{x}^i_t)$ for updating particle weights will be defined in section \ref{sec:candidate_particles}.
 
-After selecting the particle with the highest weight as the most likely candidate, the final resampling step simply consists of drawing a new set of $N$ particles with equal weights (normalized to one) from the distribution $p(\vec{x}_t|\vec{z}_{1:t})$ where $p(\vec{x}_t|\vec{z}_{1:t})$ is approximated by the set of weighted particles.
-
-\begin{equation}
-	p(\vec{x}_t|\vec{z}_{1:t}) \approx \sum_{i=1}^{N}{w^i_t\delta_{\vec{x}^i_t}}
-\end{equation}
-
-In AprilTrack, we model $p(\vec{x}_t|\vec{x}^i_{t-1})$ as a multivariate gaussian distribution with noise $\vec{\sigma}$ around 
-\begin{equation}
+In AprilTrack, we model $p(\vec{x}_t|\vec{x}^i_{t-1})$ as a multivariate gaussian distribution with standard deviation $\vec{\sigma}$ around a mean $\vec{\mu}$ where
+\begin{equation} \label{eq:propagation_mean}
 \vec{\mu} = 
 \begin{bmatrix} 
-	\vec{r}^i_{t-1} + \Delta t * \dot{\vec{r}}^i_{t-1} \\
+	\vec{r}^i_{t-1} + \Delta t \dot{\vec{r}}^i_{t-1} \\
  	\dot{\vec{r}}^i_{t-1} \\
@@
 
-where $\Delta t$ is the elapsed time and $\dot{\vec{q}}^i_{t-1}$ is the quaternion representation of $\Delta t * \vec{\omega}^i_{k-1}$ as described in [cite].
-We also use the \emph{noise quaternion} from [cite] to apply the gaussian noise to the quaternion $\vec{q}^i_{k-1} \dot{\vec{q}}^i_{k-1}$.
-For simplicity, we make the assumption that the input images have been taken at equidistant time intervals and incorporate $\Delta t$ into $\vec{\sigma}$.
+In equation \eqref{eq:propagation_mean}, $\Delta t$ is the elapsed time and $\dot{\vec{q}}^i_{t-1}$ is the quaternion representation of $\Delta t  \vec{\omega}^i_{k-1}$, as described in \cite{KalmanFilter}.
+We also use the \emph{noise quaternion} formulated in \cite{KalmanFilter} to apply the gaussian noise to the quaternion $\vec{q}^i_{k-1} \dot{\vec{q}}^i_{k-1}$.
+For simplicity, we make the assumption that the input images have been taken at equidistant time intervals and incorporate $\Delta t$ into the $\vec{\omega}$ and $\dot{\vec{r}}$ components of $\vec{\sigma}$.
 
-\subsection{Evaluation of Candidate Particles}
+After completing the update prediction and update steps and selecting the particle with the highest weight as the most likely candidate, the final resampling step simply consists of drawing a new set of $N$ particles with equal weights (normalized to a total of one) from the distribution $p(\vec{x}_t|\vec{z}_{1:t})$ where $p(\vec{x}_t|\vec{z}_{1:t})$ is approximated by the set of weighted particles.
 
-The successful tracking of AprilTags in 6 DOF requires a method for evaluating the likelihood of candidate particles at arbitrary three dimensional positions and rotations.
+\begin{equation}
+	p(\vec{x}_t|\vec{z}_{1:t}) \approx \sum_{i=1}^{N}{w^i_t\delta_{\vec{x}^i_t}}
+\end{equation}
+
+\subsection{Evaluation of Candidate Particles} \label{sec:candidate_particles}
+
+The tracking of AprilTags in 6 DOF requires a method for evaluating the likelihood of candidate particles at arbitrary three dimensional positions and rotations, or $p(\vec{z}_t|\vec{x}^i_t)$.
 To this effect, the proposed method uses homographies to generate image \emph{patches}, which are unprojected representations of an AprilTag for a particular pose and image.
 
-Given the function $I(\begin{bmatrix} \lambda x, & \lambda y, & \lambda \end{bmatrix}^\intercal )$ representing the intensity of pixel at a particular $(x,y)$ of an image, a \emph{patch} in the image can be defined by the function
+Given the function $I(\begin{bmatrix} \lambda x, & \lambda y, & \lambda \end{bmatrix}^\intercal )$ representing the intensity of pixel at a particular $(x,y)$ of an image, a patch in the image can be defined by the function
 
 \begin{equation}
-P(x_p,y_p| \vec{H}, I ) = I(\vec{H} * \begin{bmatrix} x_p & y_p & 1\end{bmatrix}^\intercal)
+P(x_p,y_p| \vec{H}, I ) = I(\vec{H}  \begin{bmatrix} x_p & y_p & 1\end{bmatrix}^\intercal)
 \end{equation}
@@
 
-This allows for the evaluation of candidate particles by the measuring the similarity between the particle's patch $P^i(x_p,y_p|\vec{H}^i,I)$ in the current image and the patch from the AprilTag's last successful detection, $P_{ref}(x_p,y_p|\vec{H}_{ref},I)$.
+This allows for the evaluation of candidate particle $\vec{x}^i$ by the measuring the similarity between the particle's patch $P^i(x_p,y_p|\vec{H}^i,I)$ in the current image and the patch from the AprilTag's last successful detection, $P_{ref}(x_p,y_p|\vec{H}_{ref},I_{ref})$.
 Here, $H^i$ can be calculated with
@@
 
-where $\widehat{\vec{q}}$ is the rotation matrix representation of particle's rotation quaternion, $\vec{r}$ is the particle's translation, and $\tau$ is the tag's edge length.
-The patch for the detected AprilTag uses the homography $\vec{H}_{ref}$ supplied by the AprilTag library, which computes the homography using the Direct Linear Transform algorithm [cite].
+where $\widehat{\vec{q}^i}$ is the rotation matrix representation of particle $\vec{x}^i$'s rotation quaternion $\vec{q}$, $\vec{r}$ is the particle's translation, and $\tau$ is the tag's edge length.
+The patch for the detected AprilTag uses the homography $\vec{H}_{ref}$ supplied by the AprilTag library, which computes the homography using the Direct Linear Transform algorithm \cite{AprilTag}.
 
@@
 
+As in BLUT \cite{BLUT}, the patch error $\epsilon(\vec{M}^{i}, \vec{M}_{ref})$ is used to compute the observation likelihood $p(\vec{z}_t|\vec{x}^i_t)$ to update the weights and select the best particle with
+
+\begin{equation}
+	p(\vec{z}_t|\vec{x}^i_t) = exp(-\gamma \epsilon(\vec{M}^i, \vec{M}_{ref}))
+\end{equation}
+
+where the parameter $\gamma$ allows for more or less selective resampling.
+
 \begin{figure}[b]
@@
 	\label{fig:error}
-	\includegraphics[width=\textwidth,height=8cm]{Average_Error}
+	\includegraphics[width=\textwidth,height=9cm]{Average_Error}
 	\caption{The average corner error between the hand-marked tag location and the tracked location.
@@
 
-\begin{figure}[b]
-	\centering
-	\includegraphics[width=7cm,height=4.75cm]{Particles2}
-	\caption{The distribution of particles and their weights with respect to the X,Y position of the particles.}
-\end{figure}
-
-\begin{figure}[b]
-	\centering
-	\includegraphics[width=7cm,height=4.75cm]{RefPatches}
-	\caption{The reference patches of various tags (by ID) in a multi-Tag tracking scenario.}
-\end{figure}
-
-As in [cite] the patch error $\epsilon(\vec{M}^{i}, \vec{M}_{ref})$ is used to compute the observation likelihood $p(\vec{z}_t|\vec{x}_t)$ to update the weights and select the best particle with
-
-\begin{equation}
-	p(\vec{z}_t|\vec{x}^i_t) = exp(-\gamma \epsilon(\vec{M}^i, \vec{M}_{ref}))
-\end{equation}
-
-where the parameter $\gamma$ allows for more or less selective resampling.
-
 \subsection{Multi-Tag Scenarios}
 
-To adapt this method to multi-tag scenarios, several extra parameters were introduced.
+To adapt this method to multi-tag scenarios, several extra parameters have to be introduced.
 In order to jointly track multiple tags, it is assumed that the relative poses of the tags are known.
@@
 where $\vec{X}_t$ is a coordinate relative to the tag's reference frame, $\vec{X}_c$ is relative to the camera reference frame, and $\vec{q}^i$ and $\vec{r}^i$ are the position and orientation of particle $\vec{x}^i_t$.
-In the simple case of jointly moving tags, the $\vec{H}^i$ from equation \ref{eq:particle_homography} can then be rewritten as
+In the simple case of jointly moving tags, the $\vec{H}^i$ from equation \eqref{eq:particle_homography} can then be rewritten as
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=7cm,height=4.75cm]{RefPatches}
+	\caption{The reference patches of various tags (by ID) in a multi-Tag tracking scenario.}
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=7cm,height=4.75cm]{Particles2}
+	\caption{The distribution of particles and their weights with respect to the X,Y position of the particles.}
+\end{figure}
+
 \section{Results and Discussion}
@@
 The algorithm was also run on an ego-motion multi-tag sequence which will be qualitatively discussed.
-
-\footnote{Videos of the sequences and the tracker's performance are available online at https://www.youtube.com/watch?v=BmZLKFBkcxU, https://www.youtube.com/watch?v=ebWaxO7JhKE, and }
+\footnote{Videos of the sequences and the tracker's performance are available online at https://www.youtube.com/watch?v=kV8z3st9rlk and https://www.youtube.com/watch?v= }
 
@@
 
-When evaluated on the single-tag sequence, the algorithm ran with $N=6000$ at around $0.5$ fps on an i7 Haswell Intel CPU.
+When evaluated on the single-tag sequence, the algorithm ran with $N=3000$ at around $1$ fps on an i7 Haswell Intel CPU.
 In experiments with fewer particles, it was found that the algorithm could track a tag well with as few as $1000$ particles with around $5$ fps.
-With a more optimized C++ implementation, the framerate could likely be brought up to $10-15$ fps, making the algorithm (run with less agressive parameters) a possible candidate for real time scenarios.
-Currently, the MATLAB implementation is intended for offline post-processing of a sequence.
+With a more optimized C++ implementation, the framerate could perhaps be brought up to $10-15$ fps, making the algorithm (run with less agressive parameters) a possible candidate for real-time scenarios.
+Currently, the MATLAB implementation is intended for offline post-processing of a sequence and so real-time performance was not a focus of this research.
 
-The pixel corner error hovers at around 10 pixels for most of the sequence.
-The tracker loses the tag momentarily during a quick twist at around frame 200 but is able to recover rapidly without any manual intervention.
+As shown in figure \ref{fig:error}, the pixel corner error hovers at around 10 pixels for most of the sequence.
+The tracker loses the tag momentarily around frames 30 and 200 but is able to recover rapidly without any manual intervention.
 Much of the error observed throughout the sequence is the result of disagreement on the boundary of a blurred tag rather than a systematic issue with the tracker itself.
@@
 Rather, the majority of the reprojection error arises from error in the relative tag positions.
+This opens up new potential avenues of research, such as tag-based bundle adjustment, to alleviate these issues.
 
@@
 
+\bibliographystyle{unsrt}
+\bibliography{paper_bib}{}
+
 \end{document}

### 2017-01-17 TagDetector.m

@@ -36,17 +36,9 @@ classdef TagDetector < TagSource
                 H = homography_solve(pin, corners);
                 [R, T] = homography_extract_pose(this.K, H);
                 r.id
-                
-                H_0 =  [0.5417    0.0623    0.6447;
-                        0.0138    0.5204    0.1269;
-                        0.0001    0.0001    0.0009];
 
-                pix_center = H(:,3) ./ H(3,3);
-                
-                T_rel = inv(H_0) * pix_center;
-                T_rel = T_rel ./ T_rel(3);
-                T_rel * 1000
-                %rot = rotm_to_quat(R')';
+                rot = rotm_to_quat(R')'
+                T
                 
                 %}
                 

### 2017-01-17 algorithm.m

@@ -1,8 +1,7 @@
-function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, record)
+function [ imgs, corners, poses ] = algorithm(tracker, detector, images, record)
     imgs = {};
     corners = [];
-    pos = [];
-    rot = [];
+    poses = [];
     
     disp('Initializing figures');
     
@@ -62,8 +61,10 @@ function [ imgs, corners, pos, rot ] = algorithm(tracker, detector, images, reco
         % tags array
         tags = project_tags(tracker.tagParams.K, tags);
         
-        pos = [pos; tags{1}.state'];
-        corners = [corners; reshape(tags{1}.corners', [1, 8])];
+        if record && length(tags) > 0
+            poses = [poses; tags{1}.state'];
+            corners = [corners; reshape(tags{1}.corners', [1, 8])];
+        end
         
         %drawTags(detector_tags, 'symbol', 'x');
         %drawTags(reproj_tags, 'symbol', 'o');

### 2017-01-17 config/movcam1_config.m

@@ -19,15 +19,15 @@ function [tracker, detector, frames] = movcam1_config(skip_rate)
 
     % Add a motion model
     mmParams(1).err_discard_threshold = 1.1;
-    mmParams(1).num_particles = 5000;
-    mmParams(1).process_noise = [0.0 0.0 0.0 ...
-                                 0.1 0.1 0.1 ...
-                                 0.0 0.0 0.0 ...
-                                 0.0 0.0 0.0];
-%    mmParams(1).process_noise = [0.02 0.02 0.02 ...
+    mmParams(1).num_particles = 3000;
+%    mmParams(1).process_noise = [0.05 0.05 0.05 ...
+%                                 0.05 0.05 0.05 ...
 %                                 0.02 0.02 0.02 ...
-%                                 0.01 0.01 0.01 ...
-%                                 0.01 0.01 0.01];
+%                                 0.02 0.02 0.02];
+    mmParams(1).process_noise = [0.02 0.02 0.02 ...
+                                 0.02 0.02 0.02 ...
+                                 0.01 0.01 0.01 ...
+                                 0.01 0.01 0.01];
 
     % For measurement error --> weight conversion
     % where weight = e^(-lambda * measurement)
@@ -37,9 +37,13 @@ function [tracker, detector, frames] = movcam1_config(skip_rate)
 
     model.loadTags('../data/movcam_tags.txt');
     
-    initial = [-0.6899; -0.1881; 3.4893; ...
-               0.1540;  -0.0023; 0.9792; 0.1319; ...
-               0; 0; 0;  0; 0; 0];
+    %initial = [0.424; 0.7353; -3.4589; ...
+    %             0.9919; -0.0018; 0.1246; 0.0231; ...
+    %               0; 0; 0;  0; 0; 0];
+    
+    initial = [0.4673; 0.4104; -3.6941; ...
+                0.9986; -0.0519; 0.0000; -0.0028; ...
+                0; 0; 0;  0; 0; 0];
     
     model.initializeParticlesTo(initial);
     

### 2017-01-17 config/test_video_config.m

@@ -19,18 +19,10 @@ function [tracker, detector, frames] = test_video_config()
     % Add a motion model
     mmParams(1).err_discard_threshold = 0.9;
     mmParams(1).num_particles = 6000;
-    mmParams(1).process_noise = [0.01 0.01 0.01 ...
-                                 0.05 0.05 0.05 ...
+    mmParams(1).process_noise = [0.005 0.005 0.005 ...
+                                 0.03 0.03 0.03 ...
                                  0.01 0.01 0.01 ...
-                                 0.01 0.01 0.01];
-
-    % noise = 1 / (k + alpha * w) * process_noise * randn
-    % Allows for particles of higher/lower weight to have
-    % more noise in the propagation step
-
-    % ATM, we resample after every pass, so don't bother...
-    mmParams(1).k = 1;
-    mmParams(1).alpha = 0;
+                                 0.03 0.03 0.03];
 
     % For measurement error --> weight conversion
     % where weight = e^(-lambda * measurement)

### 2017-01-17 run.m

@@ -1,11 +1,10 @@
-function [ imgs, corners, pos, rot ] = run(config, shouldRecord)
+function [ imgs, corners, poses, rot ] = run(config, shouldRecord)
     tracker = [];
     detector = [];
     frames = [];
     imgs = {};
     corners = [];
-    pos = [];
-    rot = [];
+    poses = [];
     
     % The parameter argument represents the number of frames to skip
     % so zero means always stay on the first frame
@@ -24,8 +23,8 @@ function [ imgs, corners, pos, rot ] = run(config, shouldRecord)
     end
     
     if (exist('shouldRecord', 'var') && shouldRecord)
-        [imgs, corners, pos, rot] = algorithm(tracker, detector, frames, true);
+        [imgs, corners, poses] = algorithm(tracker, detector, frames, true);
     else
-        [imgs, corners, pos, rot] = algorithm(tracker, detector, frames, false);
+        [imgs, corners, poses] = algorithm(tracker, detector, frames, false);
     end
 end

### 2017-01-17 paper/paper.tex
Made quick fix to paper

@@
 	\caption{The average corner error between the hand-marked tag location and the tracked location.
-Much of the error from frames 0-200 comes from discrepancy between the tracker and the ground truth on the boundaries of blurred tags.
-At around frame 200, the tracker loses the tag due to a sudden rotation but is able to recover without needing a hard reset.}
+Much of the error comes from discrepancy between the tracker and the ground truth on the boundaries of the blurred tags.
+At around frames 30 and 200, the tracker loses the tag momentarily due to a sudden rotation but is able to recover without needing to be manually reset.}
 \end{figure*}

### 2017-01-26 paper/paper.tex
Using latex rendered graph now instead of matlab graph

@@
 \usepackage{amsmath,amsfonts,amssymb}
+\usepackage{pgfplots}
+
+\pgfplotsset{compat=1.7}
 
@@
 \begin{figure*}[t]
+	\centering
 	\label{fig:error}
-	\includegraphics[width=\textwidth,height=9cm]{Average_Error}
+	\begin{tikzpicture}[trim axis left]
+	\begin{axis}[
+		width=\textwidth,
+		height=9cm,
+		title= {Average Corner Error on the Single-Tag Sequence},
+		xlabel = {Time (Frame Number)},
+		ylabel = {Average Corner Error (Pixels)},
+		xmin=0, xmax=250,
+		ymin=0, ymax=80,
+		xtick={0,50,100,150,200,250},
+		ytick={0,10,20,30,40,50,60, 70}
+	]
+	\addplot [
+		color=blue,
+		mark=none,
+	] table {corner_error.dat};
+\end{axis}
+\end{tikzpicture}
 	\caption{The average corner error between the hand-marked tag location and the tracked location.
@@
 	\centering
-	\includegraphics[width=7cm,height=4.75cm]{Particles2}
+	\includegraphics[width=8cm,height=6cm]{Particles2}
 	\caption{The distribution of particles and their weights with respect to the X,Y position of the particles.}

### 2017-01-30 paper/paper.tex
Trying to get bold stuff fixed

@@
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
+\documentclass[letterpaper, 10 pt]{IEEEconf}
 
-\usepackage{cite}
+%\makeatletter
+%\let\NAT@parse\undefined
+%\makeatother
+
 \usepackage{graphicx}
@@
 \usepackage{pgfplots}
+
+%\usepackage{natbib}
+%\usepackage{hyperref}
+%\usepackage{caption}
+%\usepackage{url}
 
@@
 
-\renewcommand{\vec}[1]{\boldsymbol{#1}}
+\renewcommand{\vec}[1]{#1}
 
 \renewcommand{\eqref}[1]{\textup{{\normalfont(\ref{#1}}\normalfont)}}
+
+%\newcommand{\citep}[1]{\cite{#1}}
+%\newcommand{\url}[1]{#1}
 
@@
 As the particles are drawn from the proposal distribution $q(\vec{x})$, the sample weight $w^i$ must take into account any discrepancy between $p(\vec{x})$ and $q(\vec{x})$ so that $w^i_t \propto p(\vec{x}^i_{1:t}|\vec{z}_{1:t})/q(\vec{x}^i_{1:t}|\vec{z}_{1:t})$.
+From this observation we note that therefore:
+
+\begin{equation*}
+	w^i_{t} \propto \frac{p(\vec{x}^i_{0:t}|\vec{z}_{1:t})}{q(\vec{x}^i_{0:t}|\vec{z}_{1:t}) }
+\end{equation*}
 
@@
 The algorithm was also run on an ego-motion multi-tag sequence which will be qualitatively discussed.
-\footnote{Videos of the sequences and the tracker's performance are available online at https://www.youtube.com/watch?v=kV8z3st9rlk and https://www.youtube.com/watch?v= }
+\footnote{Videos of the sequences and the tracker's performance are available online at https://www.youtube.com/watch?v=kV8z3st9rlk and https://www.youtube.com/watch?v=F5cFTLBlJ1s }
 
@@
 
-\bibliographystyle{unsrt}
+\bibliographystyle{abbrv}
 \bibliography{paper_bib}{}

### 2017-01-31 paper/paper.tex
Added full particle filter derivation

--- 2017-01-26 paper/paper.tex
@@
 
-\usepackage{cite}
+\makeatletter
+\let\NAT@parse\undefined
+\makeatother
+
 \usepackage{graphicx}
@@
 \usepackage{pgfplots}
+
+\usepackage{natbib}
+\usepackage{hyperref}
 
@@
 
-Visual tracking of fiducial markers (or simply fiducials) plays an important role in many robotics and computer vision related fields and applications, including augmented and virtual reality \cite{AugmentedReality}, visual odometry/SLAM \cite{TagSLAM}, and object tracking.
+Visual tracking of fiducial markers (or simply fiducials) plays an important role in many robotics and computer vision related fields and applications, including augmented and virtual reality \citep{AugmentedReality}, visual odometry/SLAM \citep{TagSLAM}, and object tracking.
 However, as many of these applications involve the use of fiducials in structured indoors environments where illumination is often limited, long exposure times to compensate for the reduced lighting produce blurred images, especially when taken from a mobile platform such as a quadrotor or a handheld device.
@@
 
- The majority of previous literature in fiducial detection has been focused single-frame fiducial detection \cite{RUNETag, BlurResilient, AprilTag}.
+ The majority of previous literature in fiducial detection has been focused single-frame fiducial detection \citep{RUNETag, BlurResilient, AprilTag}.
 A common assumption in these algorithms is that the image being processed shows little motion blur, an expectation which breaks down in empirical situations.
-For instance, the widely-used AprilTag marker \cite{AprilTag} and its successor, the AprilTag 2 \cite{AprilTag_2}, rely on strong gradients between the tag's background and the surrounding white border to detect the tag.
+For instance, the widely-used AprilTag marker \citep{AprilTag} and its successor, the AprilTag 2 \citep{AprilTag_2}, rely on strong gradients between the tag's background and the surrounding white border to detect the tag.
 As shown in figure \ref{fig:blurred_tag}, oftentimes such algorithms fail to detect tags that have been blurred, either missing them entirely or mistakenly discarding them when eliminating false positives through some encoded fiducial ID (for AprilTags this consists of a 6x6 binary grid in the center of the tag which is used to verify the validity of the detection).
@@
 Meghshyam et al.
-\cite{BlurResilient} have developed ringed markers to ensure that a region of the tag will always be perpendicular to the blur direction.
+\citep{BlurResilient} have developed ringed markers to ensure that a region of the tag will always be perpendicular to the blur direction.
 However, this approach, while resistent to linear blur, does not take into account more complex blurs (e.g resulting from the tag pivoting around the vertical axis).
-In addition, the increased resilience to blur comes at the cost of localization accuracy as only the center point of the ringed tag can be reliably determined \cite{BlurResilient}, whereas with other fiducials, such as the AprilTag \cite{AprilTag}, full pose estimation in 6 degrees of freedom (DOF) can be done based on just a single tag.
+In addition, the increased resilience to blur comes at the cost of localization accuracy as only the center point of the ringed tag can be reliably determined \citep{BlurResilient}, whereas with other fiducials, such as the AprilTag \citep{AprilTag}, full pose estimation in 6 degrees of freedom (DOF) can be done based on just a single tag.
 Another significant disadvantage of the ringed marker from Meghshyam et al.
-is the limited number of tag ID's, as their proposed encoding method allows for a maximum of 4 distinct tags \cite{BlurResilient}.
+is the limited number of tag ID's, as their proposed encoding method allows for a maximum of 4 distinct tags \citep{BlurResilient}.
 
@@
 Previous work involving the tracking of blurred objects includes the Blur-driven Tracker (BLUT) framework proposed by Wu et al.
-\cite{BLUT} for tracking motion-blurred targets.
+\citep{BLUT} for tracking motion-blurred targets.
 BLUT, a particle-filter based tracker, works on the observation that blur strength and direction can be used estimate an object's velocity and provide valuable information for successfully tracking a blurred object.
@@
 
-The novel method proposed in this paper augments the AprilTag detection algorithm originally designed by Olson \cite{AprilTag} and improved by Wang et al \cite{AprilTag_2} with a particle-filter based algorithm to reliably track an AprilTag through an arbitrary sequence of images, providing valuable fiducial-based position information when existing methods would be unable to do so.
+The novel method proposed in this paper augments the AprilTag detection algorithm originally designed by Olson \citep{AprilTag} and improved by Wang et al \citep{AprilTag_2} with a particle-filter based algorithm to reliably track an AprilTag through an arbitrary sequence of images, providing valuable fiducial-based position information when existing methods would be unable to do so.
 In addition, the proposed algorithm is extended to incorporate the joint motion of an arbitrary number of tags to increase the accuracy of the tracked orientation and position.
@@
 The AprilTrack algorithm is similar in concept to the BLUT tracker by Wu et al.
-\cite{BLUT} in that it uses a particle filter to integrate all available information about a marker's position.
+\citep{BLUT} in that it uses a particle filter to integrate all available information about a marker's position.
 Overall, the description of the algorithm can be broken down into three parts: one describing the particle filter, another on the method for evaluating candidate particles, and one describing the modifications made to simulatenously track multiple tags.
@@
 \textbf{Optimal filtering.} The prediction step can be formulated in a Bayesian setting as the computation of the prior distribution $p(\vec{x}_{t}|\vec{z}_{t-1})$ from the filtering distribution $p(\vec{x}_{t-1}|\vec{z}_{1:t-1})$ where $\vec{z}_t$ represents the measurement (appearance of a fiducial) taken at time $t$.
-As in \cite{ParticleNotes}:
+As in \citep{ParticleNotes}:
 
@@
 As the model $p(\vec{x}_t|\vec{x}_{t-1})$ is Markovian, the current measurement is conditionally independent of previous measurements given $\vec{x}_t$, meaning that $p(\vec{z}_{t}|\vec{x}_t, \vec{z}_{1:t-1})$ can be written as $p(\vec{z}_t|\vec{x}_t)$.
-$p(\vec{z}_{1:t})$ can also be expressed as $p(\vec{z}_t|\vec{z}_{1:t-1})p(\vec{z}_{1:t-1})$ resulting in the final update equation \cite{ParticleNotes}
+$p(\vec{z}_{1:t})$ can also be expressed as $p(\vec{z}_t|\vec{z}_{1:t-1})p(\vec{z}_{1:t-1})$ resulting in the final update equation \citep{ParticleNotes}
 \begin{equation} \label{eq:update}
@@
 
-\textbf{Particle Filters.} In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods.
-A particle filter consists of a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ drawn from the proposal distribution $q(\vec{x}^i)$ where each sample is called a \emph{particle}.
-This allows for the filtering distribution $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$ to be approximated by (as described in \cite{ParticleNotes})
+\textbf{Particle Filters.}\footnote{A more in-depth discussion of particle filters can found in .
+This } In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods.
+A particle filter consists of a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ drawn from the proposal distribution $q(\vec{x}^i)$ and weighted so as to take into account any discrepancy between $q(\vec{x})$ and $p(\vec{x})$ in that $w^i_t \propto p(\vec{x}^i_{0:t}|\vec{z}_{1:t})/q(\vec{x}^i_{0:t}|\vec{z}_{1:t})$.
+
+As with the optimal filtering scenario, the prior distribution (the distribution of particles at the current timestep before taking into account the measurement) is calculated in the propagation step.
+In a particle filter, this simply amounts to drawing a new set of particles based using
+
+\begin{equation} \label{eq:propagate_particles}
+	\vec{x}^i_t \sim q(\vec{x}_t|\vec{x}^i_{t-1}, \vec{z}_t)
+\end{equation}
+
+where $\sim$ denotes sampling from the state transition distribution $q(\vec{x}_t|\vec{x}^i_{t-1}, \vec{z}_t)$.
+
+Given the prior distribution, the particle weights must then be updated to reflect the target distribution $p(x)$. As previously noted:
+\begin{equation} \label{eq:weight_value}
+	w^i_{t} \propto \frac{p(\vec{x}^i_{0:t}|\vec{z}_{1:t})}{q(\vec{x}^i_{0:t}|\vec{z}_{1:t}) }
+\end{equation}
+We assume the proposal distribution $q(\vec{x})$ can be factored as
+\begin{equation*}
+	q(\vec{x}^i_{0:t}|\vec{z}_{1:t}) = q(\vec{x}^i_t|\vec{x}_{0:t-1})q(\vec{x}^i_{0:t-1}|\vec{z}_{1:t-1})
+\end{equation*}
+Since $p(\vec{x}^i_{0:t}|\vec{z}_{1:t})$ can be rewritten (in a similar manner to the derivation of \eqref{eq:update} with Bayes' law) as
+\begin{equation*}
+	p(\vec{x}^i_{0:t}|\vec{z}_{1:t}) = 
+	\frac{p(\vec{z}_t|\vec{x}^i_t)p(\vec{x}_t|\vec{x}^i_{0:t-1}, \vec{z}_{1:t-1})p(\vec{x}^i_{0:t-1}|\vec{z}_{1:t-1})}{p(\vec{z}_t|\vec{z}_{1:t-1})}
+\end{equation*}
+the equation from \eqref{eq:weight_value} can then be rewritten recursively in terms of $w^i_{t-1}$ with
+
+\begin{equation} \label{eq:update_weights}
+	w^i_t \propto w^i_{t-1} \frac{p(\vec{z}_t|\vec{x}^i_t)p(\vec{x}^i_t|\vec{x}^i_{t-1})}{q(\vec{x}^i_t|\vec{x}^i_{t-1},\vec{z}_t)}
+\end{equation}
+
+This allows us update the weights from the previous iteration and model the posterior distribution $p(\vec{x}_{0:t}|\vec{z}_{1:t})$ with the set of weighted particles.
+The highest weighted particle is seleted as the most likely candidate.
+
+To prevent the set particles from degenerating over time into a handful of highly weighted particles, resampling is applied.
+Resampling consists of drawing a new set of $N$ particles from the discrete approximation of the filtering distribution $p(\vec{x}_t|\vec{z}_{1:t})$ with the set of weighted particles:
 
 \begin{equation}
- 	p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1}) \approx \sum_{i=1}^{N}{w^i_{t-1}\delta_{\vec{x}^i_{0:t-1}}}
+	p(\vec{x}_t|\vec{z}_{1:t}) \approx \sum_{i=1}^{N}{w^i_t\delta_{\vec{x}^i_t}}
 \end{equation}
+where the weights of the new resampled particles are then set to $1/N$.
 
- where
-\begin{equation}
-	\sum_{i=1}^{N}{w^i_t} = 1
-\end{equation}
-
-As the particles are drawn from the proposal distribution $q(\vec{x})$, the sample weight $w^i$ must take into account any discrepancy between $p(\vec{x})$ and $q(\vec{x})$ so that $w^i_t \propto p(\vec{x}^i_{1:t}|\vec{z}_{1:t})/q(\vec{x}^i_{1:t}|\vec{z}_{1:t})$.
-
-In the simplified case of a Sequential Importance Resampling algorithm, the prediction and update equations derived from \eqref{eq:predict} and \eqref{eq:update} can, in the context of a SIR particle filter, be simplified to:
+The algorithm described and derived above is the Sequential Importance Sampling (SIS) filter.
+The AprilTrack tracker uses a Sequential Importance Resampling (SIR) filter, which deviates from the SIS algorithm in that resampling is applied after every iteration and the proposal distribution $q(\vec{x}_t|\vec{x}^i_{t-1})$ is simply the state transition distribution $p(\vec{x}_t|\vec{x}^i_{t-1})$.
+This means that the propagate and update equations from \eqref{eq:propagate_particles} and \eqref{eq:update_weights} simply become:
 
@@
 \end{equation}
-
+and (with the weights normalized to a total of 1):
 \begin{equation}
@@
 \end{equation}
-where $\vec{x}^i_t \sim p(\vec{x}_t|\vec{x}^i_{t-1})$ denotes that, during the prediction step, the particles for the prior distribution at time $t$ are sampled from the known distribution $p(\vec{x}_t|\vec{x}^i_{t-1})$.
-The distribution $p(\vec{z}_t|\vec{x}^i_t)$ for updating particle weights will be defined in section \ref{sec:candidate_particles}.
 
-In AprilTrack, we model $p(\vec{x}_t|\vec{x}^i_{t-1})$ as a multivariate gaussian distribution with standard deviation $\vec{\sigma}$ around a mean $\vec{\mu}$ where
+In AprilTrack, we model $p(\vec{x}_t|\vec{x}^i_{t-1})$ as a multivariate gaussian distribution with standard deviations $\vec{\sigma}$ around a mean $\vec{\mu}$ where
 \begin{equation} \label{eq:propagation_mean}
@@
 \end{equation}
-
-In equation \eqref{eq:propagation_mean}, $\Delta t$ is the elapsed time and $\dot{\vec{q}}^i_{t-1}$ is the quaternion representation of $\Delta t  \vec{\omega}^i_{k-1}$, as described in \cite{KalmanFilter}.
-We also use the \emph{noise quaternion} formulated in \cite{KalmanFilter} to apply the gaussian noise to the quaternion $\vec{q}^i_{k-1} \dot{\vec{q}}^i_{k-1}$.
+and where $\Delta t$ is the elapsed time and $\dot{\vec{q}}^i_{t-1}$ is the quaternion representation of $\Delta t  \vec{\omega}^i_{t-1}$, as described in \citep{KalmanFilter}.
+We also use the \emph{noise quaternion} formulated in \citep{KalmanFilter} to apply the gaussian noise to the quaternion $\vec{q}^i_{t-1} \dot{\vec{q}}^i_{t-1}$.
 For simplicity, we make the assumption that the input images have been taken at equidistant time intervals and incorporate $\Delta t$ into the $\vec{\omega}$ and $\dot{\vec{r}}$ components of $\vec{\sigma}$.
-
-After completing the update prediction and update steps and selecting the particle with the highest weight as the most likely candidate, the final resampling step simply consists of drawing a new set of $N$ particles with equal weights (normalized to a total of one) from the distribution $p(\vec{x}_t|\vec{z}_{1:t})$ where $p(\vec{x}_t|\vec{z}_{1:t})$ is approximated by the set of weighted particles.
-
-\begin{equation}
-	p(\vec{x}_t|\vec{z}_{1:t}) \approx \sum_{i=1}^{N}{w^i_t\delta_{\vec{x}^i_t}}
-\end{equation}
 
@@
 
-The tracking of AprilTags in 6 DOF requires a method for evaluating the likelihood of candidate particles at arbitrary three dimensional positions and rotations, or $p(\vec{z}_t|\vec{x}^i_t)$.
+The tracking of AprilTags in 6 DOF requires a method for evaluating the likelihood of candidate particles with arbitrary three dimensional positions and rotations in order to calculate $p(\vec{z}_t|\vec{x}^i_t)$.
 To this effect, the proposed method uses homographies to generate image \emph{patches}, which are unprojected representations of an AprilTag for a particular pose and image.
@@
 where $\widehat{\vec{q}^i}$ is the rotation matrix representation of particle $\vec{x}^i$'s rotation quaternion $\vec{q}$, $\vec{r}$ is the particle's translation, and $\tau$ is the tag's edge length.
-The patch for the detected AprilTag uses the homography $\vec{H}_{ref}$ supplied by the AprilTag library, which computes the homography using the Direct Linear Transform algorithm \cite{AprilTag}.
+The patch for the detected AprilTag uses the homography $\vec{H}_{ref}$ supplied by the AprilTag library, which computes the homography using the Direct Linear Transform algorithm \citep{AprilTag}.
 
@@
 
-As in BLUT \cite{BLUT}, the patch error $\epsilon(\vec{M}^{i}, \vec{M}_{ref})$ is used to compute the observation likelihood $p(\vec{z}_t|\vec{x}^i_t)$ to update the weights and select the best particle with
+As in BLUT \citep{BLUT}, the patch error $\epsilon(\vec{M}^{i}, \vec{M}_{ref})$ is used to compute the observation likelihood $p(\vec{z}_t|\vec{x}^i_t)$ to update the weights and select the best particle with
 
@@
 	\caption{A blurred tag being tracked. The red border indicates that the tag was not detected with the AprilTag detector.}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=7cm,height=4.75cm]{RefPatches}
+	\caption{The reference patches of various tags (by ID) in a multi-Tag tracking scenario.}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=7cm,height=4.75cm]{RefPatches}
-	\caption{The reference patches of various tags (by ID) in a multi-Tag tracking scenario.}
-\end{figure}
-\begin{figure}
-	\centering
 	\includegraphics[width=8cm,height=6cm]{Particles2}
@@
 
-This paper has presented the AprilTrack algorithm augmenting the standard AprilTag detection algorithm with a novel tracker capable of tracking the tags in 6 DOF by using homographies to evaluate candidate particles.
+We have presented the AprilTrack algorithm, which augments standard AprilTag detection algorithms with a tracker capable of tracking fiducials in 6 DOF.
 Experimental results show that the tracker is capable of accurately tracking both single tags and consolidating information from multiple tags in ego-motion and joint tag-motion scenarios.

### 2017-01-31 paper/paper.tex
Fixed footnote and regenerated pdf

@@
 
-\textbf{Particle Filters.}\footnote{A more in-depth discussion of particle filters can found in .
-This } In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods.
+\textbf{Particle Filters.}\footnote{A more in-depth discussion of particle filters can found in \citep{ParticleNotes}.
+Much of the notation used in \citep{ParticleNotes} is similar to that used here.} In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods.
 A particle filter consists of a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ drawn from the proposal distribution $q(\vec{x}^i)$ and weighted so as to take into account any discrepancy between $q(\vec{x})$ and $p(\vec{x})$ in that $w^i_t \propto p(\vec{x}^i_{0:t}|\vec{z}_{1:t})/q(\vec{x}^i_{0:t}|\vec{z}_{1:t})$.

### 2017-10-09 paper/paper.tex
Fixed title
Made typo fixes to paper

@@
 
-\title{\textbf{AprilTrack: A Particle-Filter based AprilTag Tracker}}
+\title{\textbf{AprilTrack: A Particle-Filter-based AprilTag Tracker}}
 \author{Daniel Pfrommer}
@@
 
- The majority of previous literature in fiducial detection has been focused single-frame fiducial detection \citep{RUNETag, BlurResilient, AprilTag}.
-A common assumption in these algorithms is that the image being processed shows little motion blur, an expectation which breaks down in empirical situations.
+ The majority of previous literature in fiducial detection has been focused on single-frame fiducial detection \citep{RUNETag, BlurResilient, AprilTag}.
+A common assumption in these algorithms is that the image being processed shows little motion blur, an expectation which breaks down in practice.
 For instance, the widely-used AprilTag marker \citep{AprilTag} and its successor, the AprilTag 2 \citep{AprilTag_2}, rely on strong gradients between the tag's background and the surrounding white border to detect the tag.
@@
 In addition, the increased resilience to blur comes at the cost of localization accuracy as only the center point of the ringed tag can be reliably determined \citep{BlurResilient}, whereas with other fiducials, such as the AprilTag \citep{AprilTag}, full pose estimation in 6 degrees of freedom (DOF) can be done based on just a single tag.
-Another significant disadvantage of the ringed marker from Meghshyam et al.
+Another significant disadvantage of the ringed marker of Meghshyam et al.
 is the limited number of tag ID's, as their proposed encoding method allows for a maximum of 4 distinct tags \citep{BlurResilient}.
@@
 As many high-blur scenarios involve a continuous stream of images taken at equidistant time intervals (e.g a quadroter-mounted moving camera viewing tags or a stationary camera tracking an AprilTag-labeled moving object), we propose to use knowledge of a fiducial's prior location to augment the detection algorithm.
-As such, this paper presents a novel particle-filter-based tracking algorithm (AprilTrack) capable of tracking AprilTags in 6 DOF with both joint-tag motion and camera ego-motion.
+As such, this paper presents a novel particle-filter-based tracking algorithm (AprilTrack) capable of tracking AprilTags in 6 DOF with both joint tag motion and camera ego-motion.
 
@@
 \citep{BLUT} for tracking motion-blurred targets.
-BLUT, a particle-filter based tracker, works on the observation that blur strength and direction can be used estimate an object's velocity and provide valuable information for successfully tracking a blurred object.
-While capable of reliably tracking blurred objects, BLUT suffers from the disadvantage that it can do so only in two dimensions (i.e it is assumed that the object always faces the camera) and so cannot track a sequence of involving arbitrary tag rotations and translations.
+BLUT, a particle-filter based tracker, works on the observation that blur strength and direction can be used to estimate an object's velocity and provide valuable information for successfully tracking a blurred object.
+While capable of reliably tracking blurred objects, BLUT suffers from the disadvantage that it can do so only in two dimensions (i.e it is assumed that the object always faces the camera) and so cannot track a sequence involving arbitrary tag rotations and translations.
 
-The novel method proposed in this paper augments the AprilTag detection algorithm originally designed by Olson \citep{AprilTag} and improved by Wang et al \citep{AprilTag_2} with a particle-filter based algorithm to reliably track an AprilTag through an arbitrary sequence of images, providing valuable fiducial-based position information when existing methods would be unable to do so.
+The novel method proposed in this paper augments the AprilTag detection algorithm originally designed by Olson \citep{AprilTag} and improved by Wang et al \citep{AprilTag_2} with a particle-filter based algorithm to reliably track an AprilTag through an arbitrary sequence of images, providing fiducial-based position information when existing methods fail.
 In addition, the proposed algorithm is extended to incorporate the joint motion of an arbitrary number of tags to increase the accuracy of the tracked orientation and position.
@@
 \textbf{Particle Filters.}\footnote{A more in-depth discussion of particle filters can found in \citep{ParticleNotes}.
-Much of the notation used in \citep{ParticleNotes} is similar to that used here.} In this regard, particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods.
+Much of the notation used in \citep{ParticleNotes} is similar to that used here.} Particle filters attempt to approximate the posterior distribution at time $t-1$, $p(\vec{x}_{0:t-1}|\vec{z}_{1:t-1})$, using Monte-Carlo methods.
 A particle filter consists of a set of $N$ samples $\{ (\vec{x}^i_{t-1}, w^i_{t-1}) \}_{i=1}^N$ drawn from the proposal distribution $q(\vec{x}^i)$ and weighted so as to take into account any discrepancy between $q(\vec{x})$ and $p(\vec{x})$ in that $w^i_t \propto p(\vec{x}^i_{0:t}|\vec{z}_{1:t})/q(\vec{x}^i_{0:t}|\vec{z}_{1:t})$.
@@
 
-As shown in figure \ref{fig:error}, the pixel corner error hovers at around 10 pixels for most of the sequence.
+As shown in figure \ref{fig:error}, the pixel corner error is about 10 pixels for most of the sequence.
 The tracker loses the tag momentarily around frames 30 and 200 but is able to recover rapidly without any manual intervention.
@@
 We have presented the AprilTrack algorithm, which augments standard AprilTag detection algorithms with a tracker capable of tracking fiducials in 6 DOF.
-Experimental results show that the tracker is capable of accurately tracking both single tags and consolidating information from multiple tags in ego-motion and joint tag-motion scenarios.
+Experimental results show that the tracker can track both single tags and consolidating information from multiple tags in ego-motion and joint tag-motion scenarios.
 In future research we plan to investigate other possible error metrics for patch comparison and look to deploy AprilTrack in realtime scenarios.

### 2017-10-09 paper/paper.tex
Updated abstract

@@
 
-	The AprilTrack algorithm proposed in this paper seeks to bridge two existing fields of research in computer vision literature: fiducial detection and monocular object tracking.
-With a rise in the popularity of fiducials, or artificial markers, as a source of ground-truth pose estimation for visual odometry datasets and mobile robots, there has been considerable interest in designing robust and blur-resistant fiducials.
-Rather than devising a new fiducial, this work attempts to improve on existing fiducials, such as the popular AprilTag marker, by augmenting fiducial detection algorithms with a particle-filter based monocular object tracker which reliably tracks a fiducial's position and orientation in six degrees of freedom, allowing for tag-based localization in situations where current detectors would be unable to do so.
-This novel patch-based particle filter tracking algorithm, which is extended to simulatenously track multiple tags in ego-motion scenarios, is qualitatively and quantitatively evaluated on several AprilTag sequences, demonstrating that it is capable of reliably localizing blurred tags.
+	The AprilTrack algorithm proposed in this paper bridges two existing fields of research in computer vision: the detection of fiducial markers (bar code type tags), and monocular object tracking.
+Visual tags are highly popular for establishing ground truth location in visual odometry experiments and on mobile robots, but detection can be difficult under rapid camera motion and in poor lighting conditions due to motion blur.
+In contrast with other works proposing redesigned, more blur resistant tags, we use the popular AprilTag fiducials and improve detection quality by means of a particle filter based monocular object tracker.
+Our algorithm is able to reliably track a fiducial's position and orientation in six degrees of freedom, allowing for tag-based localization in situations where current detectors fail.
+This novel patch-based particle filter tracking algorithm, which is extended to simultaneously track multiple tags in ego-motion scenarios, is qualitatively and quantitatively evaluated on several AprilTag sequences, demonstrating that it is capable of reliably localizing blurred tags.
 

### 2017-10-22 paper/paper.tex
Fixed typo
Fixed links, fixed figure numbering

@@
 \begin{figure*}[t]
+
 	\centering
-	\label{fig:error}
 	\begin{tikzpicture}[trim axis left]
@@
 \end{tikzpicture}
+	\label{fig:error}
 	\caption{The average corner error between the hand-marked tag location and the tracked location.
@@
 
-The AprilTrack algorithm was primarily evluated on a manually-labeled single-tag sequence with a moving tag.
+The AprilTrack algorithm was primarily evaluated on a manually-labeled single-tag sequence with a moving tag.
 The algorithm was also run on an ego-motion multi-tag sequence which will be qualitatively discussed.
-\footnote{Videos of the sequences and the tracker's performance are available online at https://www.youtube.com/watch?v=kV8z3st9rlk and https://www.youtube.com/watch?v=F5cFTLBlJ1s }
+\footnote{Videos of the sequence and the tracker's performance are available online at \url{https://www.youtube.com/watch?v=67LREs8fs9Q} and \url{https://www.youtube.com/watch?v=Rk87aE_9p-A}.
+The code is available online at \url{https://github.com/pfrommerd/tag-tracking-matlab}}
 
@@
 In experiments with fewer particles, it was found that the algorithm could track a tag well with as few as $1000$ particles with around $5$ fps.
-With a more optimized C++ implementation, the framerate could perhaps be brought up to $10-15$ fps, making the algorithm (run with less agressive parameters) a possible candidate for real-time scenarios.
+With a more optimized C++ implementation, the framerate could perhaps be brought up to $10$ fps, making the algorithm (run with less agressive parameters) a possible candidate for real-time scenarios.
 Currently, the MATLAB implementation is intended for offline post-processing of a sequence and so real-time performance was not a focus of this research.
 
-As shown in figure \ref{fig:error}, the pixel corner error is about 10 pixels for most of the sequence.
-The tracker loses the tag momentarily around frames 30 and 200 but is able to recover rapidly without any manual intervention.
+As shown in figure 4, the pixel corner error is about 10-20 pixels for most of the sequence.
+The tracker loses the tag momentarily around frame 30 but is able to recover rapidly without any manual intervention.
 Much of the error observed throughout the sequence is the result of disagreement on the boundary of a blurred tag rather than a systematic issue with the tracker itself.

### 2018-01-01 paper/paper.tex
Updated paper

@@
 \title{\textbf{AprilTrack: A Particle-Filter-based AprilTag Tracker}}
-\author{Daniel Pfrommer}
+\author{Daniel Pfrommer \\ \footnotesize{A video overview of this research is available online at \url{https://www.youtube.com/watch?v=6cFSK7Qb0QY}}}
 \date{}
@@
 This novel patch-based particle filter tracking algorithm, which is extended to simultaneously track multiple tags in ego-motion scenarios, is qualitatively and quantitatively evaluated on several AprilTag sequences, demonstrating that it is capable of reliably localizing blurred tags.
-
 \end{abstract}
@@
 We have presented the AprilTrack algorithm, which augments standard AprilTag detection algorithms with a tracker capable of tracking fiducials in 6 DOF.
-Experimental results show that the tracker can track both single tags and consolidating information from multiple tags in ego-motion and joint tag-motion scenarios.
+Experimental results show that the tracker can track both single tags and consolidate information from multiple tags in ego-motion and joint tag-motion scenarios.
 In future research we plan to investigate other possible error metrics for patch comparison and look to deploy AprilTrack in realtime scenarios.
